---
title: "Dirichlet Multinomial Mixture Models"
date: "December 2022"
authors: "S. Pinto and D. Sajbenova"
output:
  github_document: default
  word_document: default
  pdf_document: default
  html_document: default
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Load required packages

```{r Load packages 1, echo=TRUE, message=FALSE}
library( dplyr )
library( magrittr )
library( knitr )
library( tidyverse )
library( reshape2 )
library( ggplot2 )
```

```{r Load packages 2, echo=TRUE, message=FALSE, warning=FALSE}
library( phyloseq )
library( microbiome )
library( DirichletMultinomial )
library( microViz )
library( ggalluvial )
library( patchwork )
```

## Load the data

Here we use the compositional count abundance dataset, because of the requirements of the Dirichlet multinomial mixed model.

```{r Load Data, echo=FALSE}
physeq_mOTU <- readRDS( "/Volumes/My Passport for Mac/FECBUD studie/FECBUD GitHub/Data/FECBUD_physeq_filt_countdata.rds" ) 
```

## Preprocessing of the data

```{r Phyloseq object, echo=TRUE}
# Aggregate to family level
physeq_mOTU = aggregate_taxa( physeq_mOTU, "family" )
physeq_mOTU

# Store information in a dataframe
df = abundances( physeq_mOTU )
df.count = as.matrix( t( df ))
```

## Cluster 1000 times

Note that this function takes a long time to run.

```{r Repeat 1000 times, eval=FALSE}
# all_clusters_1000 = list() # save all clusters
# output_vector_1000 = list() # save best number of clusters
# for ( i in ( 1:1000 )) {
#   set.seed( i )
#   all_clusters_one = lapply( 1:10, dmn, count = df.count, verbose = FALSE )
#   all_clusters_1000[[ i ]] = all_clusters_one
#   laplace.met = base::sapply( all_clusters_one, DirichletMultinomial::laplace )
#   output_vector_1000[[ i ]] = all_clusters_one[[ which.min(unlist( laplace.met ))]]
# }
```

## Save results

```{r Save, eval=FALSE}
#saveRDS( all_clusters_1000, "all_clusters_1000.rds" )
#saveRDS( output_vector_1000, "output_vector_1000.rds" )
```

```{r Load 1000 Results, echo=FALSE}
all_clusters.donors <- readRDS( "/Volumes/My Passport for Mac/FECBUD studie/FECBUD GitHub/Data/all_clusters_1000_donors.rds" )
output_vector.donors <- readRDS( "/Volumes/My Passport for Mac/FECBUD studie/FECBUD GitHub/Data/output_vector_1000_donors.rds" )
```

## Evaluate clusters

```{r Number of clusters, echo=TRUE}
clus_2 = numeric()
clus_3 = numeric()
clus_4 = numeric()
clus_5 = numeric()
clus_6 = numeric()
clus_7 = numeric()

for ( i in 1:1000 ) {
  if ( length(output_vector.donors[[ i ]]@mixture$Weight ) == 2 ) {
    clus_2[ i ] = output_vector.donors[[ i ]]@goodnessOfFit[ "Laplace" ]
  }
  else if ( length( output_vector.donors[[ i ]]@mixture$Weight ) == 3 ) {
    clus_3[ i ] = output_vector.donors[[ i ]]@goodnessOfFit[ "Laplace" ]
  }
  else if ( length( output_vector.donors[[ i ]]@mixture$Weight ) == 4 ) {
    clus_4[ i ] = output_vector.donors[[ i ]]@goodnessOfFit[ "Laplace" ]
  }
  else if ( length( output_vector.donors[[ i ]]@mixture$Weight ) == 5 ) {
    clus_5[ i ] = output_vector.donors[[ i ]]@goodnessOfFit[ "Laplace" ]
  }
  else if ( length( output_vector.donors[[ i ]]@mixture$Weight ) == 6 ) {
    clus_6[ i ] = output_vector.donors[[ i ]]@goodnessOfFit[ "Laplace" ]
  }
  else if ( length( output_vector.donors[[ i ]]@mixture$Weight ) == 7 ) {
    clus_7[ i ] = output_vector.donors[[ i ]]@goodnessOfFit[ "Laplace" ]
  }
}

length( na.omit( clus_2 )) # 0
length( na.omit( clus_3 )) # 0
length( na.omit( clus_4 )) # 161
length( na.omit( clus_5 )) # 39
length( na.omit( clus_6 )) # 21 
length( na.omit( clus_7 )) # 779

# AICs
min( clus_4, na.rm = TRUE ) # 23899.43
min( clus_5, na.rm = TRUE ) # 23788.53
which.min( clus_5 ) # model 479

min( clus_6, na.rm = TRUE ) # 23813.3
min( clus_7, na.rm = TRUE ) # 23883.82
```

## Model with lowest AIC

```{r cluster 479, echo=TRUE}
output.479 = output_vector.donors[[ 479 ]]
sample.479 = apply( mixture( output.479 ), 1, which.max )
```

## Bacterial importance

```{r}
# importance of bacteria in each cluster based on comparisons with 1 cluster
p0 = fitted( all_clusters.donors[[ 479 ]][[ 1 ]], scale = TRUE ) # scale by theta -- with only 1 cluster
# fitted function describes the contribution of each taxonomic group
p4 = fitted( output.479, scale = TRUE ) # three clusters
colnames( p4 ) = paste( "m", 1:5, sep="" )
meandiff = colSums( abs( p4 - as.vector( p0 )))
# The posterior mean difference between the best and single-component Dirichlet multinomial model measures how each component differs from the population average;

# sum( meandiff )

diff = rowSums( abs( p4 - as.vector( p0 ))) # fitted values of taxa from cluster 3 - cluster 1 to get difference
o = order( diff, decreasing = TRUE ) # order the taxa based on their difference from highest to lowest

cdiff = cumsum( diff[ o ]) / sum( diff )
df = head( cbind( mean = p0[ o ], p4[ o, ], diff = diff[ o ], cdiff ), 10 ) # get the taxa corresponding to order for each cluster (m1, m2, m3)
df
```
## Plot bacterial importance

```{r Plot importance, echo=TRUE, fig.width = 11}
d = melt( fitted( output.479 ))
colnames( d ) = c( "Family", "cluster", "value" )

 
# Cluster 1
d1 = subset( d, cluster == 1 ) %>%
    arrange( value ) %>%
    mutate( Family = factor( Family, levels = unique( Family ))) %>%
    filter( abs( value ) > quantile( abs( value ), 0.8 ))
p1 = ggplot( d1, aes( x = Family, y = value )) +
    geom_bar( stat = "identity", fill = "#934864" ) +
    coord_flip()
  
# Cluster 2
d2 = subset( d, cluster == 2 ) %>%
    arrange( value ) %>%
    mutate( Family = factor( Family, levels = unique( Family ))) %>%
    filter( abs( value ) > quantile( abs( value ), 0.8 ))
p2 = ggplot( d2, aes( x = Family, y = value )) +
    geom_bar( stat = "identity", fill = "#658E67") +
    coord_flip()
  
# Cluster 3
d3 = subset( d, cluster == 3 ) %>%
    arrange( value ) %>%
    mutate( Family = factor( Family, levels = unique( Family ))) %>%
    filter( abs( value ) > quantile( abs( value ), 0.8 ))
p3 = ggplot( d3, aes( x = Family, y = value )) +
    geom_bar( stat = "identity", fill = "#F3E831" ) +
    coord_flip()

# Cluster 4
d4 = subset( d, cluster == 4 ) %>%
    arrange( value ) %>%
    mutate( Family = factor( Family, levels = unique( Family ))) %>%
    filter( abs( value ) > quantile( abs( value ), 0.8 ))
p4 = ggplot( d4, aes( x = Family, y = value )) +
    geom_bar( stat = "identity", fill = "#FF980A" ) +
    coord_flip()
  
# Cluster 5
d5 = subset( d, cluster == 5 ) %>%
    arrange( value ) %>%
    mutate( Family = factor( Family, levels = unique( Family ))) %>%
    filter( abs( value ) > quantile( abs( value ), 0.8 ))
p5 = ggplot( d5, aes( x = Family, y = value )) +
    geom_bar( stat = "identity", fill = "#E41A1C" ) +
    coord_flip()

# cluster 1 - green
# cluster 2 - blue
# cluster 3 - red
# cluster 5 - purple (like cluster 4 in other results)

# cluster 4 -- donors

plot.compositions = (p1 + p2 + p3 + p4 + p5) + plot_layout( nrow = 3 ) 
plot.compositions
```


```{r}
physeq.fam = subset_samples(physeq_mOTU, subject_id != "Donor A" )
physeq.fam = subset_samples(physeq.fam, subject_id != "Donor B" )
```

```{r}
physeq.fam@sam_data$clinical_outcome_wk14 = gsub("Partial", "None", physeq.fam@sam_data$clinical_outcome_wk14)

physeq.fam@sam_data$clinical_outcome_wk14 = factor(physeq.fam@sam_data$clinical_outcome_wk14, levels = c("None", "Good"))

physeq.fam@sam_data$clinical_outcome_wk10[49:57] = "Good"
physeq.fam@sam_data$clinical_outcome_wk10[78:86] = "None"

physeq.fam@sam_data[["timepoint"]] = factor(physeq.fam@sam_data[["timepoint"]],
levels = c("Baseline", "FMT1" , "FMT2", "FMT3", "FMT4", "Week7", "Week8", "Week10", "Week14"))
```

```{r}
sample.479.nodonors = sample.479[-c(181:207)]

physeq.fam@sam_data$cluster = sample.479.nodonors
physeq.fam@sam_data$cluster = as.factor(physeq.fam@sam_data$cluster)
```

```{r}
physeq.comp = microbiome::transform(physeq.fam, "compositional")
df.fam = abundances(physeq.comp)
df.fam = t(df.fam)
df.fam = as.data.frame(df.fam)

df.count = abundances(physeq.fam)
df.count = t(df.count)
df.count = as.data.frame(df.count)
```

```{r}
df.all = data.frame(df.fam, physeq.fam@sam_data)
df.all$timepoint.new = as.factor(df.all$timepoint.new)
df.all$treated_with_donor = as.factor(df.all$treated_with_donor)
df.all$pretreatment = as.factor(df.all$pretreatment)
df.all$clinical_outcome_wk14 = as.factor(df.all$clinical_outcome_wk14)
df.all$cluster = as.factor(df.all$cluster)

df.all$timepoint.new = factor(df.all$timepoint.new,
levels = c("Baseline", "Pre-FMT" , "Post-1", "Post-2", "Post-3", "Post-4", "Week8", "Week10", "Week14"))


df.all.count = data.frame(df.count, physeq.fam@sam_data)
```

```{r}
df.sub = df.all %>% select(subject_id, timepoint.new, cluster, clinical_outcome_wk14)

df.sub$cluster = as.factor(df.sub$cluster)
df.sub$timepoint.new = factor(df.sub$timepoint.new, levels = c("Baseline", "Pre-FMT", "Post-1", "Post-2", "Post-3", "Post-4", "Week8", "Week10", "Week14"))
```

```{r}
df.sub$clinical_outcome_wk14 = recode_factor(df.sub$clinical_outcome_wk14, None = "Non-Responder", Good = "Responder")
```


```{r}
# relative abundance barplot
prev.mean = df.all %>% group_by(cluster) %>% summarize(mean_prev = mean(Prevotellaceae))

rumin.mean = df.all %>% group_by(cluster) %>% summarize(mean_rum = mean(Ruminococcaceae))

lachno.mean = df.all %>% group_by(cluster) %>% summarize(mean_lach = mean(Lachnospiraceae))

clostrfam.mean = df.all %>% group_by(cluster) %>% summarize(mean_closfam = mean(Clostridialesfam.incertaesedis))

clostr.mean = df.all %>% group_by(cluster) %>% summarize(mean_clostr = mean(Clostridiaceae))

bacter.mean = df.all %>% group_by(cluster) %>% summarize(mean_bacter = mean(Bacteroidaceae))

```

```{r}
# Load required libraries
library(RColorBrewer)
library(dplyr)

df.bar = data.frame(rumin.mean, lachno.mean, clostrfam.mean, prev.mean, clostr.mean, bacter.mean)
df.bar = df.bar[, -c(3, 5, 7, 9, 11)]

df.bar.full = data.frame(cluster = rep(c(1, 2, 3, 4, 5), each = 6), taxa = rep(c("Ruminococcaceae", "Lachnospiraceae", "Clostridiales incertae sedis", "Prevotellaceae", "Clostridiaceae", "Bacteroidaceae"), times = 5), mean = c(0.2471954, 0.16642532, 0.10544134, 0.0002961352, 0.04730162, 0.12726136,
                   0.2081486, 0.08413456, 0.16572733, 0.0730258303, 0.04190005, 0.06492400,
                   0.1278547, 0.07108660, 0.06399166, 0.4719779562, 0.02613852, 0.05173534,
                   0.2396766, 0.32386105, 0.03784585, 0.0020304569, 0.06969261, 0.15670108,
                   0.2013394, 0.03128688, 0.17734669, 0.0101422621, 0.1636677, 0.02657264))

# Add an "other" group to the data frame
df.bar.full$taxa[df.bar.full$mean < 0.05] <- "Other"

# Normalize the data within each cluster to ensure bars sum up to 1
df.bar.full <- df.bar.full %>%
  group_by(cluster) %>%
  mutate(mean = mean / sum(mean))

# Change the levels of the "taxa" factor to move "Other" to the top of the legend
df.bar.full$taxa <- factor(df.bar.full$taxa, levels = c("Other", "Bacteroidaceae", "Clostridiaceae", "Clostridiales incertae sedis", "Lachnospiraceae", "Prevotellaceae", "Ruminococcaceae"))

# Create a new color palette
new_palette <- brewer.pal(8, "Dark2")  # You can adjust the number 8 based on the number of distinct groups
mycolors <- colorRampPalette( brewer.pal( 8, "Set1" ))( 16 )
mycolors <- c("#E41A1C", "#934864", "#4AA858", "#658E67", "#FF980A", "#F3E831" ,"#AB5832" )
  
# Create the plot using the new palette and modified data frame
perc.bac <- ggplot(df.bar.full, aes(fill = taxa, y = mean, x = cluster)) +
  geom_bar(position = "stack", stat = "identity", color = "black") +
  ylab("Mean Relative Abundance") +
  xlab("Cluster") +
  theme(panel.grid.major = element_blank(), panel.grid.minor = element_blank(),
        panel.background = element_blank()) +
  labs(fill = "Family") +
  scale_fill_manual(values = mycolors)  # Use the new color palette

# Print the plot
print(perc.bac)
```
## Plot clusters per patient and per timepoint

```{r Preprocessing per sample, echo=TRUE}
physeq_mOTU.patients = subset_samples(physeq_mOTU, subject_id != "Donor A" )
physeq_mOTU.patients = subset_samples(physeq_mOTU.patients, subject_id != "Donor B" )
```

```{r Get clusters, echo=TRUE}
sample.479.nodonors = sample.479[ -c( 181:207 )]

physeq_mOTU.patients@sam_data$cluster = sample.479.nodonors
physeq_mOTU.patients@sam_data$cluster = as.factor( physeq_mOTU.patients@sam_data$cluster )
```

```{r Adjust the data, echo=TRUE}
physeq.comp = microbiome::transform( physeq_mOTU.patients, "compositional" )
df.fam = abundances( physeq.comp )
df.fam = t( df.fam )
df.fam = as.data.frame( df.fam )

df.count = abundances( physeq_mOTU.patients )
df.count = t( df.count )
df.count = as.data.frame( df.count )

# Combine abundances and sample data
df.all = data.frame( df.fam, physeq_mOTU.patients@sam_data )
df.all$timepoint.new = as.factor( df.all$timepoint.new )
df.all$treated_with_donor = as.factor( df.all$treated_with_donor )
df.all$pretreatment = as.factor( df.all$pretreatment )
df.all$clinical_outcome_wk14 = as.factor( df.all$clinical_outcome_wk14 )
df.all$cluster = as.factor( df.all$cluster )

df.all$timepoint.new = factor( df.all$timepoint.new,
levels = c( "Baseline", "Pre-FMT" , "Post-1", "Post-2", "Post-3", "Post-4", "Week8", "Week10", "Week14" ))

df.all.count = data.frame( df.count, physeq_mOTU.patients@sam_data ) 

# Select patient ID, time, cluster and clinical outcome
df.sub = df.all %>% select( subject_id, timepoint.new, cluster, clinical_outcome_wk14 )

df.sub$cluster = as.factor( df.sub$cluster )
df.sub$timepoint.new = factor( df.sub$timepoint.new, levels = c( "Baseline", "Pre-FMT", "Post-1", "Post-2", "Post-3", "Post-4", "Week8", "Week10", "Week14" ))
df.sub$clinical_outcome_wk14 = recode_factor( df.sub$clinical_outcome_wk14, None = "Non-Responder", Good = "Responder" )
```

```{r Make the plot, echo=TRUE}
# cluster 1 - green
# cluster 2 - blue
# cluster 3 - red
# cluster 5 - purple (like cluster 4 in other results)

# cluster 4 -- donors
colors = c( "#934864", "#658E67", "#F3E831", "#FF980A", "#E41A1C" ) #, "#4AA858"

clus <- ggplot( df.sub, aes( x = timepoint.new, y = subject_id, fill = cluster )) + xlab( "Timepoint" ) +
       geom_tile( color = "grey20" ) + ylab( "Subject ID" ) + labs( fill = "Cluster" ) +
  facet_grid( clinical_outcome_wk14 ~ ., space = "free_y", scales = "free_y" ) +
  theme( text = element_text( size = 20 ), panel.grid.major = element_blank(), panel.grid.minor = element_blank(), panel.border = element_blank()) + scale_fill_manual( values = colors ) 
clus
# subject 105 and 121 were both from Donor A
```

```{r Visualize clusters per clinical response, echo=TRUE}
# clinical response
df.bar.resp = df.all %>% group_by( cluster ) %>% count( clinical_outcome_wk14 ) %>% mutate( freq = n / sum( n ))

df.bar.resp$cluster = as.factor( df.bar.resp$cluster )

df.bar.resp

colors = c(  "#934864", "#658E67", "#F3E831", "#FF980A", "#E41A1C" )

bars <- ggplot( df.bar.resp, aes( fill = cluster, y = n, x = clinical_outcome_wk14 )) + geom_bar( position = "fill", stat = "identity", width = 0.7 ) + ylab( "Percentage (%)" ) + theme( text = element_text( size = 20 ), panel.grid.major = element_blank(), panel.grid.minor = element_blank(),
panel.background = element_blank()) + theme( axis.title.x = element_blank()) + labs( fill = "Cluster" ) + scale_fill_manual( values = colors ) + theme( aspect.ratio = 2/1 ) +
  theme_bw()
bars
```

## PCA plot clusters of patients

```{r PCA, echo=TRUE}
pca.1.2 = physeq_mOTU.patients %>% tax_transform( "clr", rank = "unique" ) %>% # best graph
  ord_calc( method = "PCA" ) %>%
  ord_plot(
    axes = c( 1, 2 ), 
    color = "cluster", fill = "cluster", shape = "clinical_outcome_wk14", plot_taxa = c( "Prevotellaceae", "Ruminococcaceae", "Lachnospiraceae", "Clostridialesfam.incertaesedis" ), size = 2
  ) + scale_color_manual(values = c(  "#934864", "#658E67", "#F3E831", "#FF980A", "#E41A1C" )) +
  # scale_shape_manual(values = c(
    # Baseline = "circle", FMT1 = "circle cross", FMT2 = "circle open", FMT3 = "square", FMT4 = "square cross", Week7 = "square open",
    # Week8 = "triangle", Week10 = "triangle open", Week14 = "diamond" )) +
  # ggplot2::stat_ellipse(
    # ggplot2::aes( colour = clinical_outcome_wk14 )
  #) +
  ggtitle( # title
    label = "PCA Based on Aitchison Distance",
  ) +
  theme( text = element_text( size = 20 )) +
  labs( caption = "180 samples, 93 taxa" ) # caption at bottom

pca.1.2
```

## PCA plot with donors

```{r PCA donors, echo=TRUE}
physeq_mOTU@sam_data$cluster = sample.479
physeq_mOTU@sam_data$cluster = as.factor( physeq_mOTU@sam_data$cluster )

physeq_mOTU@sam_data$type = c( rep( "Subject", 180 ), rep( "Donor A", 13 ), rep( "Donor B", 14 ))
physeq_mOTU@sam_data$type = as.factor( physeq_mOTU@sam_data$type )

pca.1.3 = physeq_mOTU %>% tax_transform( "clr", rank = "unique" ) %>% # best graph
  ord_calc( method = "PCA" ) %>%
  ord_plot(
    axes = c( 1, 2 ), 
    color = "cluster", fill = "cluster", shape = "type", plot_taxa = c( "Prevotellaceae", "Ruminococcaceae", "Lachnospiraceae", "Clostridialesfam.incertaesedis" ), size = 3
  ) + scale_color_manual(values = c(  "#934864", "#658E67", "#F3E831", "#FF980A", "#E41A1C" )) +
  # scale_shape_manual( values = c(
    # Baseline = "circle", FMT1 = "circle cross", FMT2 = "circle open", FMT3 = "square", FMT4 = "square cross", Week7 = "square open",
    # Week8 = "triangle", Week10 = "triangle open", Week14 = "diamond" )) +
  # ggplot2::stat_ellipse(
    # ggplot2::aes( colour = clinical_outcome_wk14 )
  # ) +
  ggtitle( # title
    label = "PCA Based on Aitchison Distance",
  ) +
  #labs( caption = "180 samples, 93 taxa" ) + # caption at bottom
    theme( text = element_text( size = 20 )) +
  ggplot2::stat_ellipse(
    ggplot2::aes( colour = cluster )
  ) +
  theme_bw()

pca.1.3
```

```{r Combine plots, eval=FALSE}
plot.clusters = ( clus  + labs( tag = 'A' ) + theme(legend.position = "none") | bars + labs( tag = 'B' )) / pca.1.3 + labs( tag = 'C' )+ plot_layout( nrow = 2 ) + plot_layout(guides = "collect")
plot.clusters
```

## Alluvial plot clustering
```{r Alluvial, echo=TRUE}
#install.packages("ggalluvial")
# library(ggalluvial)
# library(ggplot2)
# library(ggraph)
# library(ggdist)

physeq_mOTU.patients@sam_data$cluster = as.factor( physeq_mOTU.patients@sam_data$cluster )
physeq_mOTU.patients@sam_data$timepoint.new = factor( physeq_mOTU.patients@sam_data$timepoint.new, levels = c( "Baseline", "Pre-FMT", "Post-1", "Post-2", "Post-3", "Post-4", "Week8", "Week10", "Week14" ))

two_colors = c( "#85C1E9", "#EC7063" )

pl1 = ggplot( physeq_mOTU.patients@sam_data,
       aes( x = timepoint.new, stratum = cluster, alluvium = subject_id,
           fill = clinical_outcome_wk14, label = clinical_outcome_wk14 )) +
  ylab( "Number of Patients" ) +
  scale_fill_manual( values = two_colors ) +
  geom_flow( stat = "alluvium" ) +
  geom_stratum( color = "white" ) +
  geom_text( stat = "stratum", aes( label = after_stat( stratum )))

pl2 = pl1 + theme( axis.title.x = element_blank(),  panel.grid.major = element_blank(),
    panel.grid.minor = element_blank(),
    panel.border = element_blank(),
    panel.background = element_blank()) + guides( fill = guide_legend( title = "Response" )) 

pl2
```
```{r Combine plots 2, eval=FALSE}
plot.clusters = ( clus  + labs( tag = 'A' ) + theme(legend.position = "none") | bars + labs( tag = 'B' )) / pl2 + labs( tag = 'C' )+ plot_layout( nrow = 2 ) + plot_layout(guides = "collect")
plot.clusters
```