---
title: "Top 10 families Mixed models"
date: "December 2022"
output:
  github_document: default
  word_document: default
  pdf_document: default
  html_document: default
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Load required packages

```{r Load packages 1, echo=TRUE, message=FALSE}
library( dplyr )
library( magrittr )
library( knitr )
library( tidyverse )
```

```{r Load packages 2, echo=TRUE, message=FALSE, warning=FALSE}
library( phyloseq )
library( microbiome )
library( microViz )
library( lattice )
library( latticeExtra )
library( nlme )
library( glmmTMB )
library( splines )
library( lme4 )
library( MuMIn ) 
library( aod )
library( DHARMa )
library( jtools )
library( dotwhisker )
```

```{r Load packages 3, echo=TRUE, message=FALSE}
library( broom )
library( broom.mixed )
library( dotwhisker )
library( tidyverse )
library( tibble )
```

## Load the data

Here we use the compositional count abundance dataset, else problems with convergence occur.

```{r Load Data, echo=FALSE}
physeq_mOTU <- readRDS( "/Volumes/My Passport for Mac/FECBUD studie/FECBUD GitHub/FECBUD_microbiome/Data/FECBUD_physeq_filt_countdata.rds" ) 
```

## Preprocessing of the data

```{r Phyloseq object, echo=TRUE}
# Make compositional
physeq_mOTU = microbiome::transform(physeq_mOTU, "compositional")

# Aggregate to family level
physeq_mOTU = aggregate_taxa( physeq_mOTU, "family" )
physeq_mOTU
```

```{r Preprocessing, echo=TRUE}
# Remove donors from data
physeq_mOTU = subset_samples( physeq_mOTU, subject_id != "Donor A" )
physeq_mOTU = subset_samples( physeq_mOTU, subject_id != "Donor B" )
```

```{r Sample data, echo=TRUE}
# Sample data
sample.data = as.data.frame(as.matrix(physeq_mOTU@sam_data))
sample.data$timepoint.new = as.factor(sample.data$timepoint.new)
sample.data$timepoint.new = factor(sample.data$timepoint.new, levels = c("Baseline", "Pre-FMT" , "Post-1", "Post-2", "Post-3", "Post-4", "Week8", "Week10", "Week14"))
sample.data$timepoint.new.num = as.numeric(sample.data$timepoint.new)
sample.data$clinical_outcome_wk14 = as.factor(sample.data$clinical_outcome_wk14)
sample.data$clinical_outcome_wk14 = factor(sample.data$clinical_outcome_wk14, levels = c("None", "Good"))
sample.data$sex = as.factor(sample.data$sex)
sample.data$age = as.numeric(sample.data$age)
sample.data$subject_id = as.factor(sample.data$subject_id)
sample.data$pretreatment = as.factor(sample.data$pretreatment)
sample.data$treated_with_donor = as.factor(sample.data$treated_with_donor)

# Abundance data
abund = as.data.frame(as.matrix(t(physeq_mOTU@otu_table)))

# Combine
df = data.frame(abund, sample.data)
```

# Prevotellaceae

## Transform the data 

```{r Histogramm Prevotellaceae, echo=TRUE}
hist( df$Prevotellaceae )
df$Prevotellaceae.new = asin( sqrt( df$Prevotellaceae ))
hist( df$Prevotellaceae.new )
```

```{r Proportion Prevotellaceae, echo=TRUE}
sum( df$Prevotellaceae.new == 0 ) / length( df$Prevotellaceae.new ) # proportion of zeros
```

```{r Plot data Prevotellaceae, echo=TRUE}
# None responders
df.none = subset( df, clinical_outcome_wk14 == "None" ) # changed filter to subset

# Good responders
df.good = subset( df, clinical_outcome_wk14 == "Good" )
```

## Plot Prevotellaceae

```{r Plot Prevotellaceae, echo=TRUE, fig.width = 9}
a = xyplot( Prevotellaceae.new ~ timepoint.new | clinical_outcome_wk14, data = df.none, xlab = "Timepoint", ylab = "Relative abundance (transformed)", main="Prevotellaceae", 
           panel = function( x, y ) {
             panel.average( x, y, horizontal = FALSE, col = "#EC7063", lwd = 4 )
           },
           key = list( space = "right", 
                      lines = list( col = c( "#EC7063", "#82E0AA" ), lty = c( 1, 1 ), lwd = 2 ), 
                      text = list( c( "Non-Responders", "Responders" ))))

b = xyplot( Prevotellaceae.new ~ timepoint.new | clinical_outcome_wk14, data = df.good, xlab = "Timepoint",
           panel = function( x, y ) {
             panel.average( x, y, horizontal = FALSE, col = "#82E0AA", lwd = 4, type = "l", lty = 1 )
           })

c = xyplot( Prevotellaceae.new ~ timepoint.new | clinical_outcome_wk14, data = df.none, type = "p", col = "#EC7063" )

d = xyplot( Prevotellaceae.new ~ timepoint.new | clinical_outcome_wk14, data = df.good, type = "p", col = "#82E0AA" )

Prevotellaceae.plot = a + as.layer( b ) + as.layer( c ) + as.layer( d )
Prevotellaceae.plot
```

## Choose between the lmer and glmmTMB models Prevotellaceae

We have used time as a numeric value and added a spline at knots = 7.

```{r Mixed model Prevotellaceae, echo=TRUE}
lme.int.Prevotellaceae = lmer( Prevotellaceae.new ~ sex + age + pretreatment + treated_with_donor + clinical_outcome_wk14 * ns( timepoint.new.num, knots = 7 ) + ( 1 | subject_id ), data = df )

# lme.slope.Prevotellaceae = lmer( Prevotellaceae.new ~ sex + age + pretreatment + treated_with_donor + clinical_outcome_wk14 * ns(timepoint.new.num, knots = 7) + ( ns( timepoint.new.num, knots = 7 ) | subject_id ), data = df )

glmm.int.Prevotellaceae = glmmTMB( Prevotellaceae.new ~ sex + age + pretreatment + treated_with_donor + clinical_outcome_wk14 * ns( timepoint.new.num, knots = 7 ) + ( 1 | subject_id ), 
               family = gaussian, 
               data = df, 
               zi = ~ 1,
               REML = TRUE )

# glmm.slope.Prevotellaceae = glmmTMB( Prevotellaceae.new ~ sex + age + pretreatment + treated_with_donor + clinical_outcome_wk14 * ns( timepoint.new.num, knots = 7 ) + ( ns( timepoint.new.num, knots = 7 ) | subject_id ), 
               #family = gaussian, 
               #data = df, 
               #zi = ~ 1,
               #REML = TRUE )

AICc( lme.int.Prevotellaceae ) # lower
AICc( glmm.int.Prevotellaceae )
```

# Diagnostics Prevotellaceae

```{r Diagnostics Prevotellaceae, eval=FALSE}
lme.Prevotellaceae.diag = DHARMa::simulateResiduals( lme.int.Prevotellaceae )
plot( lme.Prevotellaceae.diag )

plotResiduals( lme.Prevotellaceae.diag, form = df$clinical_outcome_wk14 )
plotResiduals( lme.Prevotellaceae.diag, form = df$timepoint.new )
plotResiduals( lme.Prevotellaceae.diag, form = df$treated_with_donor )
plotResiduals( lme.Prevotellaceae.diag, form = df$sex ) 
plotResiduals( lme.Prevotellaceae.diag, form = df$age )
plotResiduals( lme.Prevotellaceae.diag, form = df$pretreatment )

testZeroInflation( lme.Prevotellaceae.diag )
testDispersion( lme.Prevotellaceae.diag )

output1 = recalculateResiduals( lme.Prevotellaceae.diag, group = df$timepoint.new )
testTemporalAutocorrelation( output1, time = unique( df$timepoint.new ))
```

As there were no major differences in diagnostics between LMM and ZIGMM, the LMM model was chosen due to its lower AICc value. 

```{r Results Prevotellaceae}
plot_summs( lme.int.Prevotellaceae )

wald.test( Sigma = vcov( lme.int.Prevotellaceae ), b = fixef( lme.int.Prevotellaceae ), Terms = c( 6, 9, 10 ))$result # clinical outcome wk14
wald.test( Sigma = vcov( lme.int.Prevotellaceae ), b = fixef( lme.int.Prevotellaceae ), Terms = 2 )$result # sex
wald.test( Sigma = vcov( lme.int.Prevotellaceae ), b = fixef( lme.int.Prevotellaceae ), Terms = 3 )$result # age
wald.test( Sigma = vcov( lme.int.Prevotellaceae ), b = fixef( lme.int.Prevotellaceae ), Terms = 4 )$result # pretreatment
wald.test( Sigma = vcov( lme.int.Prevotellaceae ), b = fixef( lme.int.Prevotellaceae ), Terms = 5 )$result # donor
```


# Lachnospiraceae

## Transform the data 

```{r Histogramm Lachnospiraceae, echo=TRUE}
hist( df$Lachnospiraceae )
df$Lachnospiraceae.new = asin( sqrt( df$Lachnospiraceae ))
hist( df$Lachnospiraceae.new )
```

```{r Proportion Lachnospiraceae, echo=TRUE}
sum( df$Lachnospiraceae.new == 0 ) / length( df$Lachnospiraceae.new ) # proportion of zeros
```

```{r Plot data Lachnospiraceae, echo=TRUE}
# None responders
df.none = subset( df, clinical_outcome_wk14 == "None" ) # changed filter to subset

# Good responders
df.good = subset( df, clinical_outcome_wk14 == "Good" )
```

## Plot Lachnospiraceae
 
```{r Plot Lachnospiraceae, echo=TRUE, fig.width = 9}
a = xyplot( Lachnospiraceae.new ~ timepoint.new | clinical_outcome_wk14, data = df.none, xlab = "Timepoint", ylab = "Relative abundance (transformed)", main="Lachnospiraceae", 
           panel = function( x, y ) {
             panel.average( x, y, horizontal = FALSE, col = "#EC7063", lwd = 4 )
           },
           key = list( space = "right", 
                      lines = list( col = c( "#EC7063", "#82E0AA" ), lty = c( 1, 1 ), lwd = 2 ), 
                      text = list( c( "Non-Responders", "Responders" ))))

b = xyplot( Lachnospiraceae.new ~ timepoint.new | clinical_outcome_wk14, data = df.good, xlab = "Timepoint",
           panel = function( x, y ) {
             panel.average( x, y, horizontal = FALSE, col = "#82E0AA", lwd = 4, type = "l", lty = 1 )
           })

c = xyplot( Lachnospiraceae.new ~ timepoint.new | clinical_outcome_wk14, data = df.none, type = "p", col = "#EC7063" )

d = xyplot( Lachnospiraceae.new ~ timepoint.new | clinical_outcome_wk14, data = df.good, type = "p", col = "#82E0AA" )

Lachnospiraceae.plot = a + as.layer( b ) + as.layer( c ) + as.layer( d )
Lachnospiraceae.plot
```

## Choose between the lmer and glmmTMB models Lachnospiraceae

We have used time as a numeric value and added a spline at knots = 7.

```{r Mixed model Lachnospiraceae, echo=TRUE}
lme.int.Lachnospiraceae = lmer( Lachnospiraceae.new~ sex + age + pretreatment + treated_with_donor + clinical_outcome_wk14 * ns( timepoint.new.num, knots = 7 ) + ( 1 | subject_id ), data = df )

# lme.slope.Lachnospiraceaee = lmer( Lachnospiraceae.new ~ sex + age + pretreatment + treated_with_donor + clinical_outcome_wk14 * ns( timepoint.new.num, knots = 7 ) + ( ns( timepoint.new.num, knots = 7 ) | subject_id ),data = df )

glmm.int.Lachnospiraceae = glmmTMB( Lachnospiraceae.new ~ sex + age + pretreatment + treated_with_donor + clinical_outcome_wk14 * ns( timepoint.new.num, knots = 7 ) + ( 1 | subject_id ), 
               family = gaussian, 
               data = df, 
               zi = ~ 1,
               REML = TRUE )

# glmm.slope.Lachnospiraceae = glmmTMB( Lachnospiraceae.new ~ sex + age + pretreatment + treated_with_donor + clinical_outcome_wk14 * ns( timepoint.new.num, knots = 7 ) + ( ns( timepoint.new.num, knots = 7 ) | subject_id ), 
               #family = gaussian, 
               #data = df, 
               #zi = ~ 1,
               #REML = TRUE )

AICc( lme.int.Lachnospiraceae ) # lower
AICc( glmm.int.Lachnospiraceae )
```

# Diagnostics Lachnospiraceae

```{r Diagnostics Lachnospiraceae, eval=FALSE}
lme.Lachnospiraceae.diag = DHARMa::simulateResiduals( lme.int.Lachnospiraceae )
plot( lme.Lachnospiraceae.diag )

plotResiduals( lme.Lachnospiraceae.diag, form = df$clinical_outcome_wk14 )
plotResiduals( lme.Lachnospiraceae.diag, form = df$timepoint.new )
plotResiduals( lme.Lachnospiraceae.diag, form = df$treated_with_donor )
plotResiduals( lme.Lachnospiraceae.diag, form = df$sex ) 
plotResiduals( lme.Lachnospiraceae.diag, form = df$age )
plotResiduals( lme.Lachnospiraceae.diag, form = df$pretreatment )

testZeroInflation( lme.Lachnospiraceae.diag )
testDispersion( lme.Lachnospiraceae.diag )

output1 = recalculateResiduals( lme.Lachnospiraceae.diag, group = df$timepoint.new )
testTemporalAutocorrelation( output1, time = unique( df$timepoint.new ))
```

LMM was chosen due to its lower AICc value compared to ZIGMM as well as its adequate diagnostics.

```{r Results Lachnospiraceae}
plot_summs( lme.int.Lachnospiraceae )

wald.test( Sigma = vcov( lme.int.Lachnospiraceae ), b = fixef( lme.int.Lachnospiraceae ), Terms = c( 6, 9, 10 ))$result # clinical outcome wk14
wald.test( Sigma = vcov( lme.int.Lachnospiraceae ), b = fixef( lme.int.Lachnospiraceae ), Terms = 2 )$result # sex
wald.test( Sigma = vcov( lme.int.Lachnospiraceae ), b = fixef( lme.int.Lachnospiraceae ), Terms = 3 )$result # age
wald.test( Sigma = vcov( lme.int.Lachnospiraceae ), b = fixef( lme.int.Lachnospiraceae ), Terms = 4 )$result # pretreatment
wald.test( Sigma = vcov( lme.int.Lachnospiraceae ), b = fixef( lme.int.Lachnospiraceae ), Terms = 5 )$result # donor
```

# Ruminococcaceae

## Transform the data 

```{r Histogramm Ruminococcaceae, echo=TRUE}
hist( df$Ruminococcaceae )
# No transformation needed
```

```{r Proportion Ruminococcaceae, echo=TRUE}
sum( df$Ruminococcaceae == 0 ) / length( df$Ruminococcaceae ) # proportion of zeros
```

```{r Plot data Ruminococcaceae, echo=TRUE}
# None responders
df.none = subset( df, clinical_outcome_wk14 == "None" ) # changed filter to subset

# Good responders
df.good = subset( df, clinical_outcome_wk14 == "Good" )
```

## Plot Ruminococcaceae
 
```{r Plot Ruminococcaceae, echo=TRUE, fig.width = 9}
a = xyplot( Ruminococcaceae ~ timepoint.new | clinical_outcome_wk14, data = df.none, xlab = "Timepoint", ylab = "Relative abundance (transformed)", main="Ruminococcaceae", 
           panel = function( x, y ) {
             panel.average( x, y, horizontal = FALSE, col = "#EC7063", lwd = 4 )
           },
           key = list( space = "right", 
                      lines = list( col = c( "#EC7063", "#82E0AA" ), lty = c( 1, 1 ), lwd = 2 ), 
                      text = list( c( "Non-Responders", "Responders" ))))

b = xyplot( Ruminococcaceae ~ timepoint.new | clinical_outcome_wk14, data = df.good, xlab = "Timepoint",
           panel = function( x, y ) {
             panel.average( x, y, horizontal = FALSE, col = "#82E0AA", lwd = 4, type = "l", lty = 1 )
           })

c = xyplot( Ruminococcaceae ~ timepoint.new | clinical_outcome_wk14, data = df.none, type = "p", col = "#EC7063" )

d = xyplot( Ruminococcaceae ~ timepoint.new | clinical_outcome_wk14, data = df.good, type = "p", col = "#82E0AA" )

Ruminococcaceae.plot = a + as.layer( b ) + as.layer( c ) + as.layer( d )
Ruminococcaceae.plot
```

## Choose between the lmer and glmmTMB models Ruminococcaceae

We have used time as a numeric value and added a spline at knots = 7.

```{r Mixed model Ruminococcaceae, echo=TRUE}
lme.int.Ruminococcaceae = lme( Ruminococcaceae ~ sex + age + pretreatment + treated_with_donor + clinical_outcome_wk14 * ns( timepoint.new.num, knots = 7 ), 
               data = df, 
               random = ~ 1 | subject_id )

# lme.slope.Ruminococcaceae = lme( Ruminococcaceae ~ sex + age + pretreatment + treated_with_donor + clinical_outcome_wk14 * ns( timepoint.new.num, knots = 7 ), 
               #data = df, 
               #random = ~ ns( timepoint.new.num, knots = 7 ) | subject_id )

summary( lme.int.Ruminococcaceae )
```

# Diagnostics Ruminococcaceae

```{r Diagnostics Ruminococcaceae, eval=FALSE}
lmer.int.Ruminococcaceae = lmer( Ruminococcaceae ~ sex + age + pretreatment + treated_with_donor + clinical_outcome_wk14 * ns( timepoint.new.num, knots = 7 ) + ( 1 | subject_id ),
               data = df )

lmer.Ruminococcaceae.diag = DHARMa::simulateResiduals( lmer.int.Ruminococcaceae )
plot( lmer.Ruminococcaceae.diag )

plotResiduals( lmer.Ruminococcaceae.diag, form = df$clinical_outcome_wk14 )
plotResiduals( lmer.Ruminococcaceae.diag, form = df$timepoint.new )
plotResiduals( lmer.Ruminococcaceae.diag, form = df$treated_with_donor )
plotResiduals( lmer.Ruminococcaceae.diag, form = df$sex ) 
plotResiduals( lmer.Ruminococcaceae.diag, form = df$age )
plotResiduals( lmer.Ruminococcaceae.diag, form = df$pretreatment )

testZeroInflation( lmer.Ruminococcaceae.diag )
testDispersion( lmer.Ruminococcaceae.diag )

output1 = recalculateResiduals( lmer.Ruminococcaceae.diag, group = df$timepoint.new )
testTemporalAutocorrelation( output1, time = unique( df$timepoint.new ))
```

LMM was chosen due to its lower AICc value and adequate diagnostics. 

```{r Results Ruminococcaceae}
plot_summs( lme.int.Ruminococcaceae )

wald.test( Sigma = vcov( lme.int.Ruminococcaceae ), b = fixef( lme.int.Ruminococcaceae ), Terms = c( 6, 9, 10 ))$result # clinical outcome wk14
wald.test( Sigma = vcov( lme.int.Ruminococcaceae ), b = fixef( lme.int.Ruminococcaceae ), Terms = 2 )$result # sex
wald.test( Sigma = vcov( lme.int.Ruminococcaceae ), b = fixef( lme.int.Ruminococcaceae ), Terms = 3 )$result # age
wald.test( Sigma = vcov( lme.int.Ruminococcaceae ), b = fixef( lme.int.Ruminococcaceae ), Terms = 4 )$result # pretreatment
wald.test( Sigma = vcov( lme.int.Ruminococcaceae ), b = fixef( lme.int.Ruminococcaceae ), Terms = 5 )$result # donor
```

# Oscillospiraceae

## Transform the data 

```{r Histogramm Oscillospiraceae, echo=TRUE}
hist( df$Oscillospiraceae )
df$Oscillospiraceae.new = asin( sqrt( df$Oscillospiraceae ))
hist( df$Oscillospiraceae.new )
```

```{r Proportion Oscillospiraceae, echo=TRUE}
sum( df$Oscillospiraceae.new == 0 ) / length( df$Oscillospiraceae.new ) # proportion of zeros
```

```{r Plot data Oscillospiraceae, echo=TRUE}
# None responders
df.none = subset( df, clinical_outcome_wk14 == "None" ) # changed filter to subset

# Good responders
df.good = subset( df, clinical_outcome_wk14 == "Good" )
```

## Plot Oscillospiraceae
 
```{r Plot Oscillospiraceae, echo=TRUE, fig.width = 9}
a = xyplot( Oscillospiraceae.new ~ timepoint.new | clinical_outcome_wk14, data = df.none, xlab = "Timepoint", ylab = "Relative abundance (transformed)", main="Oscillospiraceae", 
           panel = function( x, y ) {
             panel.average( x, y, horizontal = FALSE, col = "#EC7063", lwd = 4 )
           },
           key = list( space = "right", 
                      lines = list( col = c( "#EC7063", "#82E0AA" ), lty = c( 1, 1 ), lwd = 2 ), 
                      text = list( c( "Non-Responders", "Responders" ))))

b = xyplot( Oscillospiraceae.new ~ timepoint.new | clinical_outcome_wk14, data = df.good, xlab = "Timepoint",
           panel = function( x, y ) {
             panel.average( x, y, horizontal = FALSE, col = "#82E0AA", lwd = 4, type = "l", lty = 1 )
           })

c = xyplot( Oscillospiraceae.new ~ timepoint.new | clinical_outcome_wk14, data = df.none, type = "p", col = "#EC7063" )

d = xyplot( Oscillospiraceae.new ~ timepoint.new | clinical_outcome_wk14, data = df.good, type = "p", col = "#82E0AA" )

Oscillospiraceae.plot = a + as.layer( b ) + as.layer( c ) + as.layer( d )
Oscillospiraceae.plot
```

## Choose between the lmer and glmmTMB models Oscillospiraceae

We have used time as a numeric value and added a spline at knots = 7.

```{r Mixed model Oscillospiraceae, echo=TRUE}
lme.int.Oscillospiraceae = lme( Oscillospiraceae.new ~ sex + age + pretreatment + treated_with_donor + clinical_outcome_wk14 * ns( timepoint.new.num, knots = 7), 
               data = df, 
               random = ~ 1 | subject_id )

# lme.slope.Oscillospiraceae = lme( Oscillospiraceae.new ~ sex + age + pretreatment + treated_with_donor + clinical_outcome_wk14 * ns( timepoint.new.num, knots = 7), 
               #data = df, 
               #random = ~ ns( timepoint.new.num, knots = 7 ) | subject_id )

glmm.int.Oscillospiraceae = glmmTMB( Oscillospiraceae.new ~ sex + age + pretreatment + treated_with_donor + clinical_outcome_wk14 + ns( timepoint.new.num, knots = 7 ) + clinical_outcome_wk14 * ns( timepoint.new.num, knots = 7 ) + ( 1 | subject_id ), 
               family = gaussian, 
               data = df, 
               zi= ~ 1 )

#glmm.slope.Oscillospiraceae. = glmmTMB( Oscillospiraceae.new ~ sex + age + pretreatment + treated_with_donor + clinical_outcome_wk14 + ns( timepoint.new.num, knots = 7 ) + clinical_outcome_wk14 * ns( timepoint.new.num, knots = 7 ) + ( ns( timepoint.new.num, knots = 7 ) | subject_id ), 
               #family = gaussian, 
               #data = df, 
               #zi= ~ 1 )

AICc( lme.int.Oscillospiraceae )
AICc( glmm.int.Oscillospiraceae ) # lower
```

# Diagnostics Oscillospiraceae

```{r Diagnostics Oscillospiraceae, eval=FALSE}
lmer.int.Oscillospiraceae = lmer(Oscillospiraceae.new ~ sex + age + pretreatment + treated_with_donor + clinical_outcome_wk14 * ns(timepoint.new.num, knots = 7) + (1 | subject_id), data = df)

lme.Oscillospiraceae.diag = DHARMa::simulateResiduals( lmer.int.Oscillospiraceae )
plot( lme.Oscillospiraceae.diag )

plotResiduals( lme.Oscillospiraceae.diag, form = df$clinical_outcome_wk14 )
plotResiduals( lme.Oscillospiraceae.diag, form = df$timepoint.new )
plotResiduals( lme.Oscillospiraceae.diag, form = df$treated_with_donor )
plotResiduals( lme.Oscillospiraceae.diag, form = df$sex ) 
plotResiduals( lme.Oscillospiraceae.diag, form = df$age )
plotResiduals( lme.Oscillospiraceae.diag, form = df$pretreatment )

testZeroInflation( lme.Oscillospiraceae.diag )
testDispersion( lme.Oscillospiraceae.diag )

output1 = recalculateResiduals( lme.Oscillospiraceae.diag, group = df$timepoint.new )
testTemporalAutocorrelation( output1, time = unique( df$timepoint.new ))
```

LMM was chosen due to better diagnostics than ZIGMM despite ZIGMM having a lower AICc value. 

```{r Results Oscillospiraceae}
plot_summs( lme.int.Oscillospiraceae )

wald.test( Sigma = vcov( lme.int.Oscillospiraceae ), b = fixef( lme.int.Oscillospiraceae ), Terms = c( 6, 9, 10 ))$result # clinical outcome wk14
wald.test( Sigma = vcov( lme.int.Oscillospiraceae ), b = fixef( lme.int.Oscillospiraceae ), Terms = 2 )$result # sex
wald.test( Sigma = vcov( lme.int.Oscillospiraceae ), b = fixef( lme.int.Oscillospiraceae ), Terms = 3 )$result # age
wald.test( Sigma = vcov( lme.int.Oscillospiraceae ), b = fixef( lme.int.Oscillospiraceae ), Terms = 4 )$result # pretreatment
wald.test( Sigma = vcov( lme.int.Oscillospiraceae ), b = fixef( lme.int.Oscillospiraceae ), Terms = 5 )$result # donor
```

# Eggerthellaceae

## Transform the data 

```{r Histogramm Eggerthellaceae, echo=TRUE}
hist( df$Eggerthellaceae )
df$Eggerthellaceae.new = asin( sqrt( df$Eggerthellaceae ))
hist( df$Eggerthellaceae.new )
```

```{r Proportion Eggerthellaceae, echo=TRUE}
sum( df$Eggerthellaceae.new == 0 ) / length( df$Eggerthellaceae.new ) # proportion of zeros
```

```{r Plot data Eggerthellaceae, echo=TRUE}
# None responders
df.none = subset( df, clinical_outcome_wk14 == "None" ) # changed filter to subset

# Good responders
df.good = subset( df, clinical_outcome_wk14 == "Good" )
```

## Plot Eggerthellaceae
 
```{r Plot Eggerthellaceae, echo=TRUE, fig.width = 9}
a = xyplot( Eggerthellaceae.new ~ timepoint.new | clinical_outcome_wk14, data = df.none, xlab = "Timepoint", ylab = "Relative abundance (transformed)", main="Eggerthellaceae", 
           panel = function( x, y ) {
             panel.average( x, y, horizontal = FALSE, col = "#EC7063", lwd = 4 )
           },
           key = list( space = "right", 
                      lines = list( col = c( "#EC7063", "#82E0AA" ), lty = c( 1, 1 ), lwd = 2 ), 
                      text = list( c( "Non-Responders", "Responders" ))))

b = xyplot( Eggerthellaceae.new ~ timepoint.new | clinical_outcome_wk14, data = df.good, xlab = "Timepoint",
           panel = function( x, y ) {
             panel.average( x, y, horizontal = FALSE, col = "#82E0AA", lwd = 4, type = "l", lty = 1 )
           })

c = xyplot( Eggerthellaceae.new ~ timepoint.new | clinical_outcome_wk14, data = df.none, type = "p", col = "#EC7063" )

d = xyplot( Eggerthellaceae.new ~ timepoint.new | clinical_outcome_wk14, data = df.good, type = "p", col = "#82E0AA" )

Eggerthellaceae.plot = a + as.layer( b ) + as.layer( c ) + as.layer( d )
Eggerthellaceae.plot
```

## Choose between the lmer and glmmTMB models Eggerthellaceae

We have used time as a numeric value and added a spline at knots = 7.

```{r Mixed model Eggerthellaceae, echo=TRUE}
lme.int.Eggerthellaceae = lme( Eggerthellaceae.new ~ sex + age + pretreatment + treated_with_donor + clinical_outcome_wk14 * ns( timepoint.new.num, knots = 7 ), 
               data = df, 
               random = ~ 1 | subject_id )

# lme.slope.Eggerthellaceae = lme( Eggerthellaceae.new ~ sex + age + pretreatment + treated_with_donor + clinical_outcome_wk14 * ns( timepoint.new.num, knots = 7 ), 
               # data = df, 
               # random = ~ ns( timepoint.new.num, knots = 7 ) | subject_id )

glmm.int.Eggerthellaceae = glmmTMB( Eggerthellaceae.new ~ sex + age + pretreatment + treated_with_donor + clinical_outcome_wk14 * ns( timepoint.new.num, knots = 7 ) + ( 1 | subject_id ), 
               family = gaussian, 
               data = df, 
               zi = ~ 1,
               REML = TRUE )

# glmm.slope.Eggerthellaceae = glmmTMB( Eggerthellaceae.new ~ sex + age + pretreatment + treated_with_donor + clinical_outcome_wk14 * ns( timepoint.new.num, knots = 7) + ( ns( timepoint.new, knots = 7 ) | subject_id ), 
               # family = gaussian, 
               # data = df, 
               # zi = ~ 1,
               # REML = TRUE )

AICc( lme.int.Eggerthellaceae ) # lower
AICc( glmm.int.Eggerthellaceae )
```

# Diagnostics Eggerthellaceae

```{r Diagnostics Eggerthellaceae, eval=FALSE}
lmer.int.Eggerthellaceae = lmer(Eggerthellaceae.new ~ sex + age + pretreatment + treated_with_donor + clinical_outcome_wk14 * ns(timepoint.new.num, knots = 7) + (1 | subject_id), data = df)

lme.Eggerthellaceae.diag = DHARMa::simulateResiduals( lmer.int.Eggerthellaceae )
plot( lme.Eggerthellaceae.diag )

plotResiduals( lme.Eggerthellaceae.diag, form = df$clinical_outcome_wk14 )
plotResiduals( lme.Eggerthellaceae.diag, form = df$timepoint.new )
plotResiduals( lme.Eggerthellaceae.diag, form = df$treated_with_donor )
plotResiduals( lme.Eggerthellaceae.diag, form = df$sex ) 
plotResiduals( lme.Eggerthellaceae.diag, form = df$age )
plotResiduals( lme.Eggerthellaceae.diag, form = df$pretreatment )

testZeroInflation( lme.Eggerthellaceae.diag )
testDispersion( lme.Eggerthellaceae.diag )

output1 = recalculateResiduals( lme.Eggerthellaceae.diag, group = df$timepoint.new )
testTemporalAutocorrelation( output1, time = unique( df$timepoint.new ))
```

LMM was chosen due to a lower AICc and adequate diagnostics. 

```{r Results Eggerthellaceae}
plot_summs( lme.int.Eggerthellaceae )

wald.test( Sigma = vcov( lme.int.Eggerthellaceae ), b = fixef( lme.int.Eggerthellaceae ), Terms = c( 6, 9, 10 ))$result # clinical outcome wk14
wald.test( Sigma = vcov( lme.int.Eggerthellaceae ), b = fixef( lme.int.Eggerthellaceae ), Terms = 2 )$result # sex
wald.test( Sigma = vcov( lme.int.Eggerthellaceae ), b = fixef( lme.int.Eggerthellaceae ), Terms = 3 )$result # age
wald.test( Sigma = vcov( lme.int.Eggerthellaceae ), b = fixef( lme.int.Eggerthellaceae ), Terms = 4 )$result # pretreatment
wald.test( Sigma = vcov( lme.int.Eggerthellaceae ), b = fixef( lme.int.Eggerthellaceae ), Terms = 5 )$result # donor
```

# Coefficients Plot - Comparison

```{r Coefficients plot data, echo=TRUE}
# Prevotellaceae
prevo = broom.mixed::tidy( lme.int.Prevotellaceae, conf.int = TRUE )
prevo = prevo[ -c( 11:12 ), -c( 7:8 )]
model = "Prevotellaceae"
prevo = data.frame( model = model, prevo )
prevo = as_tibble( prevo )

# Lachnospiraceae
lachno = broom.mixed::tidy( lme.int.Lachnospiraceae, conf.int = TRUE )
lachno = lachno[ -c( 11:12 ), -c( 7:8 )]
model = "Lachnospiraceae"
lachno = data.frame( model = model, lachno )
lachno = as_tibble( lachno )

# Ruminococcaceae
rumino = broom.mixed::tidy( lme.int.Ruminococcaceae, conf.int = TRUE )
rumino = rumino[ -c( 11:12 ), -c( 6,8 )]
model = "Ruminococcaceae"
rumino = data.frame( model = model, rumino )
rumino = as_tibble( rumino )

# Oscillospiraceae
oscillo = broom.mixed::tidy( lme.int.Oscillospiraceae, conf.int = TRUE )
oscillo = oscillo[ -c( 11:12 ), -c( 6,8 )]
model = "Oscillospiraceae"
oscillo = data.frame( model = model, oscillo )
oscillo = as_tibble( oscillo )

# Eggerthellaceae
eggerth = broom.mixed::tidy( lme.int.Eggerthellaceae, conf.int = TRUE )
eggerth = eggerth[ -c( 11:12 ), -c( 6,8 )]
model = "Eggerthellaceae"
eggerth = data.frame( model = model, eggerth )
eggerth = as_tibble( eggerth )
```


```{r Coefficient Plot, echo=TRUE, fig.width = 9, fig.height = 9}
# combine
common <- intersect( colnames( prevo ), colnames( rumino ))
all_bacteria = rbind( prevo[ common ], lachno[ common ], rumino[ common ], oscillo[ common ], eggerth[ common ])

# pdf( "coefficientsplot.pdf", width = 15, height = 11 )
# dwplot( all_bacteria, dodge_size = 0.8, dot_args = list( size = 3 ),
#   whisker_args = list( size = 2 )) + xlab( "Coefficient Estimate" ) + geom_vline( xintercept = 0, colour = "grey60", linetype = 2 ) + theme_bw( base_size = 4 ) + theme( legend.title = element_blank(), text = element_text( size = 20 ))

dwplot( all_bacteria ) + xlab( "Coefficient Estimate" ) + geom_vline( xintercept = 0, colour = "grey60", linetype = 2 ) + theme_bw() + theme( legend.title = element_blank())
# dev.off()
```
