---
title: "Top 10 families Mixed models"
date: "December 2022"
output:
  github_document: default
  word_document: default
  pdf_document: default
  html_document: default
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Load required packages

```{r Load packages 1, echo=TRUE, message=FALSE}
library( dplyr )
library( magrittr )
library( knitr )
library( tidyverse )
```

```{r Load packages 2, echo=TRUE, message=FALSE, warning=FALSE}
library( phyloseq )
library( microbiome )
library( microViz )
library( lattice )
library( latticeExtra )
library( nlme )
library( glmmTMB )
library( splines )
library( lme4 )
library( MuMIn ) 
library( aod )
library( DHARMa )
library( jtools )
library( dotwhisker )
```

```{r Load packages 3, echo=TRUE, message=FALSE}
library( broom )
library( broom.mixed )
library( dotwhisker )
library( tidyverse )
library( tibble )
```

## Load the data

Here we use the compositional count abundance dataset, else problems with convergence occur.

```{r Load Data, echo=FALSE}
physeq_mOTU <- readRDS( "/Volumes/My Passport for Mac/FECBUD studie/FECBUD GitHub/FECBUD_microbiome/Data/FECBUD_physeq_filt_countdata.rds" ) 
```

## Preprocessing of the data

```{r Phyloseq object, echo=TRUE}
# Make compositional
physeq_mOTU = microbiome::transform(physeq_mOTU, "compositional")

# Aggregate to family level
physeq_mOTU = aggregate_taxa( physeq_mOTU, "family" )
physeq_mOTU
```

```{r Preprocessing, echo=TRUE}
# Remove donors from data
physeq_mOTU = subset_samples( physeq_mOTU, subject_id != "Donor A" )
physeq_mOTU = subset_samples( physeq_mOTU, subject_id != "Donor B" )
```

```{r Sample data, echo=TRUE}
# Sample data
sample.data = as.data.frame(as.matrix(physeq_mOTU@sam_data))
sample.data$timepoint.new = as.factor(sample.data$timepoint.new)
sample.data$timepoint.new = factor(sample.data$timepoint.new, levels = c("Baseline", "Pre-FMT" , "Post-1", "Post-2", "Post-3", "Post-4", "Week8", "Week10", "Week14"))
sample.data$timepoint.new.num = as.numeric(sample.data$timepoint.new)
sample.data$clinical_outcome_wk14 = as.factor(sample.data$clinical_outcome_wk14)
sample.data$clinical_outcome_wk14 = factor(sample.data$clinical_outcome_wk14, levels = c("None", "Good"))
sample.data$sex = as.factor(sample.data$sex)
sample.data$age = as.numeric(sample.data$age)
sample.data$subject_id = as.factor(sample.data$subject_id)
sample.data$pretreatment = as.factor(sample.data$pretreatment)
sample.data$treated_with_donor = as.factor(sample.data$treated_with_donor)

# Abundance data
abund = as.data.frame(as.matrix(t(physeq_mOTU@otu_table)))

# Combine
df = data.frame(abund, sample.data)
```

# Bacteroidaceae

## Transform the data 

```{r Histogramm Bacteroidaceae, echo=TRUE}
hist( df$Bacteroidaceae )
df$Bacteroidaceae.new = asin( sqrt( df$Bacteroidaceae ))
hist( df$Bacteroidaceae.new )
```

```{r Proportion Bacteroidaceae, echo=TRUE}
sum( df$Bacteroidaceae.new == 0 ) / length( df$Bacteroidaceae.new ) # proportion of zeros
```

```{r Plot data Bacteroidaceae, echo=TRUE}
# None responders
df.none = subset( df, clinical_outcome_wk14 == "None" ) # changed filter to subset

# Good responders
df.good = subset( df, clinical_outcome_wk14 == "Good" )
```

## Plot Bacteroidaceae
 
```{r Plot Bacteroidaceae, echo=TRUE, fig.width = 9}
a = xyplot( Bacteroidaceae.new ~ timepoint.new | clinical_outcome_wk14, data = df.none, xlab = "Timepoint", ylab = "Relative abundance (transformed)", main="Bacteroidaceae", 
           panel = function( x, y ) {
             panel.average( x, y, horizontal = FALSE, col = "#EC7063", lwd = 4 )
           },
           key = list( space = "right", 
                      lines = list( col = c( "#EC7063", "#82E0AA" ), lty = c( 1, 1 ), lwd = 2 ), 
                      text = list( c( "Non-Responders", "Responders" ))))

b = xyplot( Bacteroidaceae.new ~ timepoint.new | clinical_outcome_wk14, data = df.good, xlab = "Timepoint",
           panel = function( x, y ) {
             panel.average( x, y, horizontal = FALSE, col = "#82E0AA", lwd = 4, type = "l", lty = 1 )
           })

c = xyplot( Bacteroidaceae.new ~ timepoint.new | clinical_outcome_wk14, data = df.none, type = "p", col = "#EC7063" )

d = xyplot( Bacteroidaceae.new ~ timepoint.new | clinical_outcome_wk14, data = df.good, type = "p", col = "#82E0AA" )

Bacteroidaceae.plot = a + as.layer( b ) + as.layer( c ) + as.layer( d )
Bacteroidaceae.plot
```

## Choose between the lmer and glmmTMB models Bacteroidaceae

We have used time as a numeric value and added a spline at knots = 7.

```{r Mixed model Bacteroidaceae, echo=TRUE}
lme.int.Bacteroidaceae = lme4::lmer(Bacteroidaceae.new ~ sex + age + pretreatment + treated_with_donor + clinical_outcome_wk14 * ns(timepoint.new.num, knots = 7) + (1 | subject_id), data = df)

#lme.int.Bacteroidaceae = lme4::lmer(Bacteroidaceae.new ~ sex + age + pretreatment + treated_with_donor + clinical_outcome_wk14 * ns(timepoint.new.num, knots = 7) + (ns(timepoint.new.num, knots = 7) | subject_id), data = df)

glmm.int.Bacteroidaceae = glmmTMB(Bacteroidaceae.new ~ sex + age + pretreatment + treated_with_donor + clinical_outcome_wk14 * ns(timepoint.new.num, knots = 7) + (1 | subject_id), 
               family = gaussian, 
               data = df, 
               zi= ~ 1) 

#glmm.slope.Bacteroidaceae = glmmTMB(Bacteroidaceae.new ~ sex + age + pretreatment + treated_with_donor + clinical_outcome_wk14 * ns(timepoint.new.num, knots = 7) + (ns(timepoint.new.num, knots = 7) | subject_id), 
               #family = gaussian, 
               #data = df, 
               #zi= ~ 1) 

AICc(lme.int.Bacteroidaceae)
AICc(glmm.int.Bacteroidaceae) # lower
```

# Diagnostics Bacteroidaceae

```{r Diagnostics Bacteroidaceae, eval=FALSE}
lme.Bacteroidaceae.diag = DHARMa::simulateResiduals( lme.int.Bacteroidaceae )
plot( lme.Bacteroidaceae.diag )

plotResiduals( lme.Bacteroidaceae.diag, form = df$clinical_outcome_wk14 )
plotResiduals( lme.Bacteroidaceae.diag, form = df$timepoint.new )
plotResiduals( lme.Bacteroidaceae.diag, form = df$treated_with_donor )
plotResiduals( lme.Bacteroidaceae.diag, form = df$sex ) 
plotResiduals( lme.Bacteroidaceae.diag, form = df$age )
plotResiduals( lme.Bacteroidaceae.diag, form = df$pretreatment )

testZeroInflation( lme.Bacteroidaceae.diag )
testDispersion( lme.Bacteroidaceae.diag )

output1 = recalculateResiduals( lme.Bacteroidaceae.diag, group = df$timepoint.new )
testTemporalAutocorrelation( output1, time = unique( df$timepoint.new ))
```

LMM model was chosen due to slightly better diagnostics. 

```{r Results Bacteroidaceae}
plot_summs( lme.int.Bacteroidaceae )

wald.test( Sigma = vcov( lme.int.Bacteroidaceae ), b = fixef( lme.int.Bacteroidaceae ), Terms = c( 6, 9, 10 ))$result # clinical outcome wk14
wald.test( Sigma = vcov( lme.int.Bacteroidaceae ), b = fixef( lme.int.Bacteroidaceae ), Terms = 2 )$result # sex
wald.test( Sigma = vcov( lme.int.Bacteroidaceae ), b = fixef( lme.int.Bacteroidaceae ), Terms = 3 )$result # age
wald.test( Sigma = vcov( lme.int.Bacteroidaceae ), b = fixef( lme.int.Bacteroidaceae ), Terms = 4 )$result # pretreatment
wald.test( Sigma = vcov( lme.int.Bacteroidaceae ), b = fixef( lme.int.Bacteroidaceae ), Terms = 5 )$result # donor
```


# Bacteroidalesfam.incertaesedis

```{r}
sum(df$Bacteroidalesfam.incertaesedis == 0) / length(df$Bacteroidalesfam.incertaesedis)
```

```{r}
xyplot(Bacteroidalesfam.incertaesedis ~ timepoint.new | clinical_outcome_wk14, data = df, xlab = "Bacteroidalesfam.incertaesedis", ylab = "Bifidobacteriaceae", ylim = c(0, 0.05),
           panel = function(x, y) {
            panel.average(x, y, horizontal = FALSE, col = "plum", lwd = 4)
       })

hist(df$Bacteroidalesfam.incertaesedis)

df$Bacteroidalesfam.incertaesedis.new = asin(sqrt(df$Bacteroidalesfam.incertaesedis))
hist(df$Bacteroidalesfam.incertaesedis.new)

xyplot(Bacteroidalesfam.incertaesedis.new ~ timepoint.new | clinical_outcome_wk14, data = df, xlab = "Bacteroidalesfam.incertaesedis", ylab = "Bifidobacteriaceae", ylim = c(0, 0.15),
           panel = function(x, y) {
            panel.average(x, y, horizontal = FALSE, col = "plum", lwd = 4)
       })
```

```{r}
lme.int.Bacteroidalesfam.incertaesedis = lmer(Bacteroidalesfam.incertaesedis.new ~ sex + age + pretreatment + treated_with_donor + clinical_outcome_wk14 * ns(timepoint.new, knots = 7) + (1 | subject_id), data = df)

#lme.slope.Bacteroidalesfam.incertaesedis = lmer(Bacteroidalesfam.incertaesedis.new ~ sex + age + pretreatment + treated_with_donor + clinical_outcome_wk14 * ns(timepoint.new, knots = 7) + (ns(timepoint.new, knots = 7) | subject_id), data = df)

glmm.int.Bacteroidalesfam.incertaesedis = glmmTMB(Bacteroidalesfam.incertaesedis.new ~ sex + age + pretreatment + treated_with_donor + clinical_outcome_wk14 * ns(timepoint.new, knots = 7) + (1 | subject_id), 
               family = gaussian, 
               data = df, 
               zi= ~ clinical_outcome_wk14)

#glmm.slope.Bacteroidalesfam.incertaesedis = glmmTMB(Bacteroidalesfam.incertaesedis.new ~ sex + age + pretreatment + treated_with_donor + clinical_outcome_wk14 * ns(timepoint.new, knots = 7) + (ns(timepoint.new, knots = 7) | subject_id), 
               #family = gaussian, 
               #data = df, 
               #zi= ~ clinical_outcome_wk14) 

AICc(lme.int.Bacteroidalesfam.incertaesedis)
AICc(glmm.int.Bacteroidalesfam.incertaesedis) # lower
```

```{r}
# diagnostics zigmm
glmm.Bacteroidalesfam.incertaesedis.diag = DHARMa::simulateResiduals(glmm.int.Bacteroidalesfam.incertaesedis)

plot(glmm.Bacteroidalesfam.incertaesedis.diag)

plotResiduals(glmm.Bacteroidalesfam.incertaesedis.diag, form = df$clinical_outcome_wk14)
plotResiduals(glmm.Bacteroidalesfam.incertaesedis.diag, form = df$timepoint.new)
plotResiduals(glmm.Bacteroidalesfam.incertaesedis.diag, form = df$treated_with_donor)
plotResiduals(glmm.Bacteroidalesfam.incertaesedis.diag, form = df$sex)
plotResiduals(glmm.Bacteroidalesfam.incertaesedis.diag, form = df$age)
plotResiduals(glmm.Bacteroidalesfam.incertaesedis.diag, form = df$pretreatment)

testZeroInflation(glmm.Bacteroidalesfam.incertaesedis.diag)
testDispersion(glmm.Bacteroidalesfam.incertaesedis.diag)
output1 = recalculateResiduals(glmm.Bacteroidalesfam.incertaesedis.diag, group=df$timepoint.new)
testTemporalAutocorrelation(output1, time = unique(df$timepoint.new))
```

```{r}
# diagnostics lme
lme.Bacteroidalesfam.incertaesedis.diag = DHARMa::simulateResiduals(lme.int.Bacteroidalesfam.incertaesedis)

plot(lme.Bacteroidalesfam.incertaesedis.diag)

plotResiduals(lme.Bacteroidalesfam.incertaesedis.diag, form = df$clinical_outcome_wk14)
plotResiduals(lme.Bacteroidalesfam.incertaesedis.diag, form = df$timepoint.new)
plotResiduals(lme.Bacteroidalesfam.incertaesedis.diag, form = df$treated_with_donor)
plotResiduals(lme.Bacteroidalesfam.incertaesedis.diag, form = df$sex)
plotResiduals(lme.Bacteroidalesfam.incertaesedis.diag, form = df$age)
plotResiduals(lme.Bacteroidalesfam.incertaesedis.diag, form = df$pretreatment)

testZeroInflation(lme.Bacteroidalesfam.incertaesedis.diag)
testDispersion(lme.Bacteroidalesfam.incertaesedis.diag)
output1 = recalculateResiduals(lme.Bacteroidalesfam.incertaesedis.diag, group=df$timepoint.new)
testTemporalAutocorrelation(output1, time = unique(df$timepoint.new))
```

```{r}
wald.test(Sigma = vcov(glmm.int.Bacteroidalesfam.incertaesedis)[["cond"]], b = fixef(glmm.int.Bacteroidalesfam.incertaesedis)[["cond"]], Terms = c(6, 9, 10))$result

wald.test(Sigma = vcov(glmm.int.Bacteroidalesfam.incertaesedis)[["cond"]], b = fixef(glmm.int.Bacteroidalesfam.incertaesedis)[["cond"]], Terms = 2)$result # sex

wald.test(Sigma = vcov(glmm.int.Bacteroidalesfam.incertaesedis)[["cond"]], b = fixef(glmm.int.Bacteroidalesfam.incertaesedis)[["cond"]], Terms = 3)$result # age

wald.test(Sigma = vcov(glmm.int.Bacteroidalesfam.incertaesedis)[["cond"]], b = fixef(glmm.int.Bacteroidalesfam.incertaesedis)[["cond"]], Terms = 4)$result # pretreatment

wald.test(Sigma = vcov(glmm.int.Bacteroidalesfam.incertaesedis)[["cond"]], b = fixef(glmm.int.Bacteroidalesfam.incertaesedis)[["cond"]], Terms = 5)$result # donor
```

# Bifidobacteriaceae

```{r}
sum(df$Bifidobacteriaceae == 0) / length(df$Bifidobacteriaceae)
```

```{r}
xyplot(Bifidobacteriaceae ~ timepoint.new | clinical_outcome_wk14, data = df, xlab = "Timepoint", ylab = "Bifidobacteriaceae", ylim = c(0, 0.03),
           panel = function(x, y) {
            panel.average(x, y, horizontal = FALSE, col = "plum", lwd = 4)
       })

hist(df$Bifidobacteriaceae)

df$Bifidobacteriaceae.new = asin(sqrt(df$Bifidobacteriaceae))
hist(df$Bifidobacteriaceae.new)

xyplot(Bifidobacteriaceae.new ~ timepoint.new | clinical_outcome_wk14, data = df, xlab = "Timepoint", ylab = "Bifidobacteriaceae", ylim = c(0, 0.15),
           panel = function(x, y) {
            panel.average(x, y, horizontal = FALSE, col = "plum", lwd = 4)
       })
```

```{r}
lme.int.Bifidobacteriaceae = lmer(Bifidobacteriaceae.new ~ sex + age + pretreatment + treated_with_donor + clinical_outcome_wk14 * ns(timepoint.new, knots = 7) + (1 | subject_id), data = df)

#lme.slope.Bifidobacteriaceae = lmer(Bifidobacteriaceae.new ~ sex + age + pretreatment + treated_with_donor + clinical_outcome_wk14 * ns(timepoint.new, knots = 7) + (ns(timepoint.new, knots = 7) | subject_id), data = df)

glmm.int.Bifidobacteriaceae = glmmTMB(Bifidobacteriaceae.new ~ sex + age + pretreatment + treated_with_donor + clinical_outcome_wk14 * ns(timepoint.new, knots = 7) + (1 | subject_id), 
               family = gaussian, 
               data = df, 
               zi= ~ 1)

#glmm.slope.Bifidobacteriaceae = glmmTMB(Bifidobacteriaceae.new ~ sex + age + pretreatment + treated_with_donor + clinical_outcome_wk14 * ns(timepoint.new, knots = 7) + (ns(timepoint.new, knots = 7) | subject_id), 
               #family = gaussian, 
               #data = df, 
               #zi= ~ 1) 

AICc(lme.int.Bifidobacteriaceae)
AICc(glmm.int.Bifidobacteriaceae) # lower
```

```{r}
# diagnostics zigmm
glmm.Bifidobacteriaceae.diag = DHARMa::simulateResiduals(glmm.int.Bifidobacteriaceae)

plot(glmm.Bifidobacteriaceae.diag)

plotResiduals(glmm.Bifidobacteriaceae.diag, form = df$clinical_outcome_wk14)
plotResiduals(glmm.Bifidobacteriaceae.diag, form = df$timepoint.new)
plotResiduals(glmm.Bifidobacteriaceae.diag, form = df$treated_with_donor)
plotResiduals(glmm.Bifidobacteriaceae.diag, form = df$sex)
plotResiduals(glmm.Bifidobacteriaceae.diag, form = df$age)
plotResiduals(glmm.Bifidobacteriaceae.diag, form = df$pretreatment)

testZeroInflation(glmm.Bifidobacteriaceae.diag)
testDispersion(glmm.Bifidobacteriaceae.diag)
output1 = recalculateResiduals(glmm.Bifidobacteriaceae.diag, group=df$timepoint.new)
testTemporalAutocorrelation(output1, time = unique(df$timepoint.new))
```

```{r}
# diagnostics lme
lme.Bifidobacteriaceae.diag = DHARMa::simulateResiduals(lme.int.Bifidobacteriaceae)

plot(lme.Bifidobacteriaceae.diag)

plotResiduals(lme.Bifidobacteriaceae.diag, form = df$clinical_outcome_wk14)
plotResiduals(lme.Bifidobacteriaceae.diag, form = df$timepoint.new)
plotResiduals(lme.Bifidobacteriaceae.diag, form = df$treated_with_donor)
plotResiduals(lme.Bifidobacteriaceae.diag, form = df$sex)
plotResiduals(lme.Bifidobacteriaceae.diag, form = df$age)
plotResiduals(lme.Bifidobacteriaceae.diag, form = df$pretreatment)

testZeroInflation(lme.Bifidobacteriaceae.diag)
testDispersion(lme.Bifidobacteriaceae.diag)
output1 = recalculateResiduals(lme.Bifidobacteriaceae.diag, group=df$timepoint.new)
testTemporalAutocorrelation(output1, time = unique(df$timepoint.new))
```

```{r}
wald.test(Sigma = vcov(glmm.int.Bifidobacteriaceae)[["cond"]], b = fixef(glmm.int.Bifidobacteriaceae)[["cond"]], Terms = c(6, 9, 10))$result

wald.test(Sigma = vcov(glmm.int.Bifidobacteriaceae)[["cond"]], b = fixef(glmm.int.Bifidobacteriaceae)[["cond"]], Terms = 2)$result # sex

wald.test(Sigma = vcov(glmm.int.Bifidobacteriaceae)[["cond"]], b = fixef(glmm.int.Bifidobacteriaceae)[["cond"]], Terms = 3)$result # age

wald.test(Sigma = vcov(glmm.int.Bifidobacteriaceae)[["cond"]], b = fixef(glmm.int.Bifidobacteriaceae)[["cond"]], Terms = 4)$result # pretreatment

wald.test(Sigma = vcov(glmm.int.Bifidobacteriaceae)[["cond"]], b = fixef(glmm.int.Bifidobacteriaceae)[["cond"]], Terms = 5)$result # donor
```


# Clostridiaceae

## Transform the data 

```{r Histogramm Clostridiaceae, echo=TRUE}
hist( df$Clostridiaceae )
df$Clostridiaceae.new = asin( sqrt( df$Clostridiaceae ))
hist( df$Clostridiaceae.new )
```

```{r Proportion Clostridiaceae, echo=TRUE}
sum( df$Clostridiaceae.new == 0 ) / length( df$Clostridiaceae.new ) # proportion of zeros
```

```{r Plot data Clostridiaceae, echo=TRUE}
# None responders
df.none = subset( df, clinical_outcome_wk14 == "None" ) # changed filter to subset

# Good responders
df.good = subset( df, clinical_outcome_wk14 == "Good" )
```

## Plot Clostridiaceae
 
```{r Plot Clostridiaceae, echo=TRUE, fig.width = 9}
a = xyplot( Clostridiaceae.new ~ timepoint.new | clinical_outcome_wk14, data = df.none, xlab = "Timepoint", ylab = "Relative abundance (transformed)", main="Clostridiaceae", 
           panel = function( x, y ) {
             panel.average( x, y, horizontal = FALSE, col = "#EC7063", lwd = 4 )
           },
           key = list( space = "right", 
                      lines = list( col = c( "#EC7063", "#82E0AA" ), lty = c( 1, 1 ), lwd = 2 ), 
                      text = list( c( "Non-Responders", "Responders" ))))

b = xyplot( Clostridiaceae.new ~ timepoint.new | clinical_outcome_wk14, data = df.good, xlab = "Timepoint",
           panel = function( x, y ) {
             panel.average( x, y, horizontal = FALSE, col = "#82E0AA", lwd = 4, type = "l", lty = 1 )
           })

c = xyplot( Clostridiaceae.new ~ timepoint.new | clinical_outcome_wk14, data = df.none, type = "p", col = "#EC7063" )

d = xyplot( Clostridiaceae.new ~ timepoint.new | clinical_outcome_wk14, data = df.good, type = "p", col = "#82E0AA" )

Clostridiaceae.plot = a + as.layer( b ) + as.layer( c ) + as.layer( d )
Clostridiaceae.plot
```

## Choose between the lmer and glmmTMB models Clostridiaceae

We have used time as a numeric value and added a spline at knots = 7.

```{r Mixed model Clostridiaceae, echo=TRUE}
lme.int.Clostridiaceae = lme( Clostridiaceae.new~ sex + age + pretreatment + treated_with_donor + clinical_outcome_wk14 * ns( timepoint.new.num, knots = 7 ), 
               data = df, 
               random = ~ 1 | subject_id )

# lme.slope.Clostridiaceae = lme( Clostridiaceae.new ~ sex + age + pretreatment + treated_with_donor + clinical_outcome_wk14 * ns( timepoint.new.num, knots = 7), 
              # data = df, 
              # random = ~ ns( timepoint.new.num, knots = 7 ) | subject_id )

glmm.int.Clostridiaceae = glmmTMB( Clostridiaceae.new ~ sex + age + pretreatment + treated_with_donor + clinical_outcome_wk14 + ns( timepoint.new.num, knots = 7 ) + clinical_outcome_wk14 * ns( timepoint.new.num, knots = 7 ) + ( 1 | subject_id ), 
               family = gaussian, 
               data = df, 
               zi= ~ 1 )

# glmm.slope.Clostridiaceae = glmmTMB( Clostridiaceae.new ~ sex + age + pretreatment + treated_with_donor + clinical_outcome_wk14 + ns( timepoint.new.num, knots = 7 ) + clinical_outcome_wk14 * ns( timepoint.new.num, knots = 7 ) + ( ns(timepoint.new.num, knots = 7 ) | subject_id ), 
               # family = gaussian, 
               # data = df, 
               # zi= ~ 1 )

AICc( lme.int.Clostridiaceae )
AICc( glmm.int.Clostridiaceae ) # lower
```

# Diagnostics Clostridiaceae

```{r Diagnostics Clostridiaceae, eval=FALSE}
glmm.Clostridiaceae.diag = DHARMa::simulateResiduals( glmm.int.Clostridiaceae )
plot( glmm.Clostridiaceae.diag )

plotResiduals( glmm.Clostridiaceae.diag, form = df$clinical_outcome_wk14 )
plotResiduals( glmm.Clostridiaceae.diag, form = df$timepoint.new )
plotResiduals( glmm.Clostridiaceae.diag, form = df$treated_with_donor )
plotResiduals( glmm.Clostridiaceae.diag, form = df$sex ) 
plotResiduals( glmm.Clostridiaceae.diag, form = df$age )
plotResiduals( glmm.Clostridiaceae.diag, form = df$pretreatment )

testZeroInflation( glmm.Clostridiaceae.diag )
testDispersion( glmm.Clostridiaceae.diag )

output1 = recalculateResiduals( glmm.Clostridiaceae.diag, group = df$timepoint.new )
testTemporalAutocorrelation( output1, time = unique( df$timepoint.new ))
```

ZIGMM model had a lower AICc and similar diagnostics to the LMM. 

```{r Results Clostridiaceae, fig.width = 9}
conf.int = data.frame( confint( glmm.int.Clostridiaceae ))
colnames( conf.int )[ 1 ] = "conf.low"
colnames( conf.int )[ 2 ] = "conf.high"
conf.int = conf.int[ -c( 11:21 ), ]

clostridfam = broom.mixed::tidy( glmm.int.Clostridiaceae )[ , -c( 1 )]
clostridfam = clostridfam[ -c( 11:23 ), ]
clostridfam = clostridfam[ , -c( 1 )]
effect = "fixed"

clostridfam = data.frame( model = "Clostridiales incertaesedis", effect = effect, clostridfam, conf.low = conf.int$conf.low, conf.high = conf.int$conf.high )
clostridfam = clostridfam[ , -c( 8 )]
clostridfam = as_tibble( clostridfam )

dwplot( clostridfam ) + xlab( "Coefficient Estimate" ) + geom_vline( xintercept = 0, colour = "grey60", linetype = 2 ) + theme_bw() + theme( legend.title = element_blank())

wald.test( Sigma = vcov( glmm.int.Clostridiaceae )[[ "cond" ]], b = fixef( glmm.int.Clostridiaceae )[[ "cond" ]], Terms = c( 6, 9, 10 ))$result

wald.test( Sigma = vcov( glmm.int.Clostridiaceae )[[ "cond" ]], b = fixef( glmm.int.Clostridiaceae )[[ "cond" ]], Terms = 2 )$result # sex

wald.test( Sigma = vcov( glmm.int.Clostridiaceae )[[ "cond" ]], b = fixef( glmm.int.Clostridiaceae )[[ "cond" ]], Terms = 3 )$result # age

wald.test( Sigma = vcov( glmm.int.Clostridiaceae )[[ "cond" ]], b = fixef( glmm.int.Clostridiaceae )[[ "cond" ]], Terms = 4 )$result # pretreatment

wald.test( Sigma = vcov( glmm.int.Clostridiaceae )[[ "cond" ]], b = fixef( glmm.int.Clostridiaceae )[[ "cond" ]], Terms = 5 )$result # donor
```


# Clostridialesfam.incertaesedis

## Transform the data 

```{r Histogramm Clostridialesfam.incertaesedis, echo=TRUE}
hist( df$Clostridialesfam.incertaesedis )
df$Clostridialesfam.incertaesedis.new = asin( sqrt( df$Clostridialesfam.incertaesedis ))
hist( df$Clostridialesfam.incertaesedis.new )
```

```{r Proportion Clostridialesfam.incertaesedis, echo=TRUE}
sum( df$Clostridialesfam.incertaesedis.new == 0 ) / length( df$Clostridialesfam.incertaesedis.new ) # proportion of zeros
```

```{r Plot data Clostridialesfam.incertaesedis, echo=TRUE}
# None responders
df.none = subset( df, clinical_outcome_wk14 == "None" ) # changed filter to subset

# Good responders
df.good = subset( df, clinical_outcome_wk14 == "Good" )
```

## Plot Clostridialesfam.incertaesedis
 
```{r Plot Clostridialesfam.incertaesedis, echo=TRUE, fig.width = 9}
a = xyplot( Clostridialesfam.incertaesedis.new ~ timepoint.new | clinical_outcome_wk14, data = df.none, xlab = "Timepoint", ylab = "Relative abundance (transformed)", main="Clostridialesfam.incertaesedis", 
           panel = function( x, y ) {
             panel.average( x, y, horizontal = FALSE, col = "#EC7063", lwd = 4 )
           },
           key = list( space = "right", 
                      lines = list( col = c( "#EC7063", "#82E0AA" ), lty = c( 1, 1 ), lwd = 2 ), 
                      text = list( c( "Non-Responders", "Responders" ))))

b = xyplot( Clostridialesfam.incertaesedis.new ~ timepoint.new | clinical_outcome_wk14, data = df.good, xlab = "Timepoint",
           panel = function( x, y ) {
             panel.average( x, y, horizontal = FALSE, col = "#82E0AA", lwd = 4, type = "l", lty = 1 )
           })

c = xyplot( Clostridialesfam.incertaesedis.new ~ timepoint.new | clinical_outcome_wk14, data = df.none, type = "p", col = "#EC7063" )

d = xyplot( Clostridialesfam.incertaesedis.new ~ timepoint.new | clinical_outcome_wk14, data = df.good, type = "p", col = "#82E0AA" )

Clostridialesfam.incertaesedis.plot = a + as.layer( b ) + as.layer( c ) + as.layer( d )
Clostridialesfam.incertaesedis.plot
```

## Choose between the lmer and glmmTMB models Clostridialesfam.incertaesedis

We have used time as a numeric value and added a spline at knots = 7.

```{r Mixed model Clostridialesfam.incertaesedis, echo=TRUE}
lme.int.Clostridialesfam.incertaesedis = lme( Clostridialesfam.incertaesedis.new~ sex + age + pretreatment + treated_with_donor + clinical_outcome_wk14 * ns( timepoint.new.num, knots = 7 ), 
               data = df, 
               random = ~ 1 | subject_id )

# lme.slope.Clostridialesfam.incertaesedis = lme( Clostridialesfam.incertaesedis.new ~ sex + age + pretreatment + treated_with_donor + clinical_outcome_wk14 * ns( timepoint.new.num, knots = 7 ), 
#                data = df, 
#                random = ~ ns( timepoint.new.num, knots = 7 ) | subject_id )

# glmm.int.Clostridialesfam.incertaesedis = glmmTMB( Clostridialesfam.incertaesedis.new ~ sex + age + pretreatment + treated_with_donor + clinical_outcome_wk14 * ns( timepoint.new.num, knots = 7 ) + ( 1 | subject_id ), 
#                family = gaussian, 
#                data = df, 
#                zi = ~ clinical_outcome_wk14 + ns( timepoint.new.num, knots = 7 ) + ( 1 | subject_id ),
#                REML = TRUE )

glmm.slope.Clostridialesfam.incertaesedis = glmmTMB( Clostridialesfam.incertaesedis.new ~ sex + age + pretreatment + treated_with_donor + clinical_outcome_wk14 * ns( timepoint.new.num, knots = 7 ) + ( ns( timepoint.new.num, knots = 7 ) | subject_id ), 
               family = gaussian, 
               data = df, 
               zi = ~ clinical_outcome_wk14 + ns(timepoint.new.num, knots = 7 ) + ( 1 | subject_id ),
               REML = TRUE )

AICc( lme.int.Clostridialesfam.incertaesedis ) 
AICc( glmm.slope.Clostridialesfam.incertaesedis ) # lower
```

# Diagnostics Clostridialesfam.incertaesedis

```{r Diagnostics Clostridialesfam.incertaesedis, eval=FALSE}
glmm.Clostridialesfam.incertaesedis.diag = DHARMa::simulateResiduals( glmm.slope.Clostridialesfam.incertaesedis )
plot( glmm.Clostridialesfam.incertaesedis.diag )

plotResiduals( glmm.Clostridialesfam.incertaesedis.diag, form = df$clinical_outcome_wk14 )
plotResiduals( glmm.Clostridialesfam.incertaesedis.diag, form = df$timepoint.new )
plotResiduals( glmm.Clostridialesfam.incertaesedis.diag, form = df$treated_with_donor )
plotResiduals( glmm.Clostridialesfam.incertaesedis.diag, form = df$sex ) 
plotResiduals( glmm.Clostridialesfam.incertaesedis.diag, form = df$age )
plotResiduals( glmm.Clostridialesfam.incertaesedis.diag, form = df$pretreatment )

testZeroInflation( glmm.Clostridialesfam.incertaesedis.diag )
testDispersion( glmm.Clostridialesfam.incertaesedis.diag )

output1 = recalculateResiduals( glmm.Clostridialesfam.incertaesedis.diag, group = df$timepoint.new )
testTemporalAutocorrelation( output1, time = unique( df$timepoint.new ))
```
As the ZIGMM showed better diagnostics, it was selected as the better model. AICc values were very similar between LMM and ZIGMM models. 

```{r Results Clostridialesfam.incertaesedis, fig.width = 9}
conf.int = data.frame( confint( glmm.slope.Clostridialesfam.incertaesedis ))
colnames( conf.int )[ 1 ] = "conf.low"
colnames( conf.int )[ 2 ] = "conf.high"
conf.int = conf.int[ -c( 11:21 ), ]

clostridfam = broom.mixed::tidy( glmm.slope.Clostridialesfam.incertaesedis )[ , -c( 1 )]
clostridfam = clostridfam[ -c( 11:23 ), ]
clostridfam = clostridfam[ , -c( 1 )]
effect = "fixed"

clostridfam = data.frame( model = "Clostridiales incertaesedis", effect = effect, clostridfam, conf.low = conf.int$conf.low, conf.high = conf.int$conf.high )
clostridfam = clostridfam[ , -c( 8 )]
clostridfam = as_tibble( clostridfam )

dwplot( clostridfam ) + xlab( "Coefficient Estimate" ) + geom_vline( xintercept = 0, colour = "grey60", linetype = 2 ) + theme_bw() + theme( legend.title = element_blank())

wald.test( Sigma = vcov( glmm.slope.Clostridialesfam.incertaesedis )[[ "cond" ]], b = fixef( glmm.slope.Clostridialesfam.incertaesedis )[[ "cond" ]], Terms = c( 6, 9, 10 ))$result

wald.test( Sigma = vcov( glmm.slope.Clostridialesfam.incertaesedis )[[ "cond" ]], b = fixef( glmm.slope.Clostridialesfam.incertaesedis )[[ "cond" ]], Terms = 2 )$result # sex

wald.test( Sigma = vcov( glmm.slope.Clostridialesfam.incertaesedis )[[ "cond" ]], b = fixef( glmm.slope.Clostridialesfam.incertaesedis )[[ "cond" ]], Terms = 3 )$result # age

wald.test( Sigma = vcov( glmm.slope.Clostridialesfam.incertaesedis )[[ "cond" ]], b = fixef( glmm.slope.Clostridialesfam.incertaesedis )[[ "cond" ]], Terms = 4 )$result # pretreatment

wald.test( Sigma = vcov( glmm.slope.Clostridialesfam.incertaesedis )[[ "cond" ]], b = fixef( glmm.slope.Clostridialesfam.incertaesedis )[[ "cond" ]], Terms = 5 )$result # donor
```

# Coriobacteriaceae

```{r}
sum(df$Coriobacteriaceae == 0) / length(df$Coriobacteriaceae) 
```

```{r}
xyplot(Coriobacteriaceae ~ timepoint.new | clinical_outcome_wk14, data = df, xlab = "Timepoint", ylab = "Coriobacteriaceae", ylim = c(0, 0.04),
           panel = function(x, y) {
            panel.average(x, y, horizontal = FALSE, col = "plum", lwd = 4)
       })

hist(df$Coriobacteriaceae)

df$Coriobacteriaceae.new = asin(sqrt(df$Coriobacteriaceae))
hist(df$Coriobacteriaceae.new)

xyplot(Coriobacteriaceae.new ~ timepoint.new | clinical_outcome_wk14, data = df, xlab = "Timepoint", ylab = "Coriobacteriaceae", ylim = c(0, 0.2),
           panel = function(x, y) {
            panel.average(x, y, horizontal = FALSE, col = "plum", lwd = 4)
       })
```

```{r}
lme.int.Coriobacteriaceae = lmer(Coriobacteriaceae.new ~ sex + age + pretreatment + treated_with_donor + clinical_outcome_wk14 * ns(timepoint.new, knots = 7) + (1 | subject_id), data = df)

#lme.slope.Coriobacteriaceae = lmer(Coriobacteriaceae.new ~ sex + age + pretreatment + treated_with_donor + clinical_outcome_wk14 * ns(timepoint.new, knots = 7) + (ns(timepoint.new, knots = 7) | subject_id), data = df)

glmm.int.Coriobacteriaceae = glmmTMB(Coriobacteriaceae.new ~ sex + age + pretreatment + treated_with_donor + clinical_outcome_wk14 * ns(timepoint.new, knots = 7) + (1 | subject_id), 
               family = gaussian, 
               data = df, 
               zi= ~ 1)

#glmm.slope.Coriobacteriaceae = glmmTMB(Coriobacteriaceae.new ~ sex + age + pretreatment + treated_with_donor + clinical_outcome_wk14 * ns(timepoint.new, knots = 7) + (ns(timepoint.new, knots = 7) | subject_id), 
               #family = gaussian, 
               #data = df, 
               #zi= ~ 1) 

AICc(lme.int.Coriobacteriaceae)
AICc(glmm.int.Coriobacteriaceae) # lower
```

```{r}
# diagnostics zigmm
glmm.Coriobacteriaceae.diag = DHARMa::simulateResiduals(glmm.int.Coriobacteriaceae)

plot(glmm.Coriobacteriaceae.diag)

plotResiduals(glmm.Coriobacteriaceae.diag, form = df$clinical_outcome_wk14)
plotResiduals(glmm.Coriobacteriaceae.diag, form = df$timepoint.new)
plotResiduals(glmm.Coriobacteriaceae.diag, form = df$treated_with_donor)
plotResiduals(glmm.Coriobacteriaceae.diag, form = df$sex)
plotResiduals(glmm.Coriobacteriaceae.diag, form = df$age)
plotResiduals(glmm.Coriobacteriaceae.diag, form = df$pretreatment)

testZeroInflation(glmm.Coriobacteriaceae.diag)
testDispersion(glmm.Coriobacteriaceae.diag)
output1 = recalculateResiduals(glmm.Coriobacteriaceae.diag, group=df$timepoint.new)
testTemporalAutocorrelation(output1, time = unique(df$timepoint.new))
```

```{r}
# diagnostics lme
lme.Coriobacteriaceae.diag = DHARMa::simulateResiduals(lme.int.Coriobacteriaceae)

plot(lme.Coriobacteriaceae.diag)

plotResiduals(lme.Coriobacteriaceae.diag, form = df$clinical_outcome_wk14)
plotResiduals(lme.Coriobacteriaceae.diag, form = df$timepoint.new)
plotResiduals(lme.Coriobacteriaceae.diag, form = df$treated_with_donor)
plotResiduals(lme.Coriobacteriaceae.diag, form = df$sex)
plotResiduals(lme.Coriobacteriaceae.diag, form = df$age)
plotResiduals(lme.Coriobacteriaceae.diag, form = df$pretreatment)

testZeroInflation(lme.Coriobacteriaceae.diag)
testDispersion(lme.Coriobacteriaceae.diag)
output1 = recalculateResiduals(lme.Coriobacteriaceae.diag, group=df$timepoint.new)
testTemporalAutocorrelation(output1, time = unique(df$timepoint.new))
```

```{r}
wald.test(Sigma = vcov(glmm.int.Coriobacteriaceae)[["cond"]], b = fixef(glmm.int.Coriobacteriaceae)[["cond"]], Terms = c(6, 9, 10))$result

wald.test(Sigma = vcov(glmm.int.Coriobacteriaceae)[["cond"]], b = fixef(glmm.int.Coriobacteriaceae)[["cond"]], Terms = 2)$result # sex

wald.test(Sigma = vcov(glmm.int.Coriobacteriaceae)[["cond"]], b = fixef(glmm.int.Coriobacteriaceae)[["cond"]], Terms = 3)$result # age

wald.test(Sigma = vcov(glmm.int.Coriobacteriaceae)[["cond"]], b = fixef(glmm.int.Coriobacteriaceae)[["cond"]], Terms = 4)$result # pretreatment

wald.test(Sigma = vcov(glmm.int.Coriobacteriaceae)[["cond"]], b = fixef(glmm.int.Coriobacteriaceae)[["cond"]], Terms = 5)$result # donor
```

# Eubacteriaceae 

```{r}
sum(df$Eubacteriaceae == 0) / length(df$Eubacteriaceae)
```

```{r}
xyplot(Eubacteriaceae ~ timepoint.new | clinical_outcome_wk14, data = df, xlab = "Timepoint", ylab = "Eubacteriaceae",
           panel = function(x, y) {
            panel.average(x, y, horizontal = FALSE, col = "plum", lwd = 4)
       })

hist(df$Eubacteriaceae)
df$Eubacteriaceae.new = asin(sqrt(df$Eubacteriaceae))
hist(df$Eubacteriaceae.new)

xyplot(Eubacteriaceae.new ~ timepoint.new | clinical_outcome_wk14, data = df, xlab = "Timepoint", ylab = "Eubacteriaceae",
           panel = function(x, y) {
            panel.average(x, y, horizontal = FALSE, col = "plum", lwd = 4)
       })
```

```{r}
library(nlme)

lme.int.Eubacteriaceae = lme(Eubacteriaceae.new ~ sex + age + pretreatment + treated_with_donor + clinical_outcome_wk14 * ns(timepoint.new, knots = 7), 
               data = df, 
               random = ~ 1 | subject_id)

lme.slope.Eubacteriaceae = lme(Eubacteriaceae.new ~ sex + age + pretreatment + treated_with_donor + clinical_outcome_wk14 * ns(timepoint.new, knots = 7), 
               data = df, 
               random = ~ ns(timepoint.new, knots = 7) | subject_id)

anova_results = anova(lme.int.Eubacteriaceae, lme.slope.Eubacteriaceae)
mixtureLRT = anova_results[["L.Ratio"]][2]
0.5 * pchisq(mixtureLRT, 1, lower.tail = FALSE) + 0.5 * pchisq(mixtureLRT, 2, lower.tail = FALSE) # use random-slopes

glmm.int.Eubacteriaceae = glmmTMB(Eubacteriaceae.new ~ sex + age + pretreatment + treated_with_donor + clinical_outcome_wk14 * ns(timepoint.new, knots = 7) + (1 | subject_id), 
               family = gaussian, 
               data = df, 
               zi = ~ 1,
               REML = TRUE)

lme.slope.Eubacteriaceae = lmer(Eubacteriaceae.new ~ sex + age + pretreatment + treated_with_donor + clinical_outcome_wk14 * ns(timepoint.new, knots = 7) + (ns(timepoint.new, knots = 7) | subject_id), data = df)

#glmm.slope.Eubacteriaceae = glmmTMB(Eubacteriaceae.new ~ sex + age + pretreatment + treated_with_donor + clinical_outcome_wk14 * ns(timepoint.new, knots = 7) + (ns(timepoint.new, knots = 7) | subject_id), 
               #family = gaussian, 
               #data = df, 
               #zi = ~ 1,
               #REML = TRUE)

AICc(lme.slope.Eubacteriaceae) # lower
AICc(glmm.int.Eubacteriaceae)
```

```{r}
# diagnostics lme
lme.Eubacteriaceae.diag = DHARMa::simulateResiduals(lme.slope.Eubacteriaceae)

plot(lme.Eubacteriaceae.diag)

plotResiduals(lme.Eubacteriaceae.diag, form = df$clinical_outcome_wk14)
plotResiduals(lme.Eubacteriaceae.diag, form = df$timepoint.new)
plotResiduals(lme.Eubacteriaceae.diag, form = df$treated_with_donor)
plotResiduals(lme.Eubacteriaceae.diag, form = df$sex)
plotResiduals(lme.Eubacteriaceae.diag, form = df$age)
plotResiduals(lme.Eubacteriaceae.diag, form = df$pretreatment)

testZeroInflation(lme.Eubacteriaceae.diag)
testDispersion(lme.Eubacteriaceae.diag)
output1 = recalculateResiduals(lme.Eubacteriaceae.diag, group=df$timepoint.new)
testTemporalAutocorrelation(output1, time = unique(df$timepoint.new))
```

```{r}
# diagnostics zigmm
glmm.Eubacteriaceae.diag = DHARMa::simulateResiduals(glmm.int.Eubacteriaceae)

plot(glmm.Eubacteriaceae.diag)

plotResiduals(glmm.Eubacteriaceae.diag, form = df$clinical_outcome_wk14)
plotResiduals(glmm.Eubacteriaceae.diag, form = df$timepoint.new)
plotResiduals(glmm.Eubacteriaceae.diag, form = df$treated_with_donor)
plotResiduals(glmm.Eubacteriaceae.diag, form = df$sex)
plotResiduals(glmm.Eubacteriaceae.diag, form = df$age)
plotResiduals(glmm.Eubacteriaceae.diag, form = df$pretreatment)

testZeroInflation(glmm.Eubacteriaceae.diag)
testDispersion(glmm.Eubacteriaceae.diag)
output1 = recalculateResiduals(glmm.Eubacteriaceae.diag, group=df$timepoint.new)
testTemporalAutocorrelation(output1, time = unique(df$timepoint.new))
```
```{r}
wald.test(Sigma = vcov(lme.slope.Eubacteriaceae), b = fixef(lme.slope.Eubacteriaceae), Terms = c(6, 9, 10))$result

wald.test(Sigma = vcov(lme.slope.Eubacteriaceae), b = fixef(lme.slope.Eubacteriaceae), Terms = 2)$result # sex

wald.test(Sigma = vcov(lme.slope.Eubacteriaceae), b = fixef(lme.slope.Eubacteriaceae), Terms = 3)$result # age

wald.test(Sigma = vcov(lme.slope.Eubacteriaceae), b = fixef(lme.slope.Eubacteriaceae), Terms = 4)$result # pretreatment

wald.test(Sigma = vcov(lme.slope.Eubacteriaceae), b = fixef(lme.slope.Eubacteriaceae), Terms = 5)$result # donor
```

# Firmicutesfam.incertaesedis

```{r}
sum(df$Firmicutesfam.incertaesedis == 0) / length(df$Firmicutesfam.incertaesedis)
```

```{r}
xyplot(Firmicutesfam.incertaesedis ~ timepoint.new | clinical_outcome_wk14, data = df, xlab = "Timepoint", ylab = "Firmicutesfam.incertaesedis", ylim = c(0, 0.04),
           panel = function(x, y) {
            panel.average(x, y, horizontal = FALSE, col = "plum", lwd = 4)
       })

hist(df$Firmicutesfam.incertaesedis)
df$Firmicutesfam.incertaesedis.new = asin(sqrt(df$Firmicutesfam.incertaesedis))
hist(df$Firmicutesfam.incertaesedis.new)

xyplot(Firmicutesfam.incertaesedis ~ timepoint.new | clinical_outcome_wk14, data = df, xlab = "Timepoint", ylab = "Firmicutesfam.incertaesedis", ylim = c(0, 0.04), 
           panel = function(x, y) {
            panel.average(x, y, horizontal = FALSE, col = "plum", lwd = 4)
       })
```

```{r}
lme.int.Firmicutesfam.incertaesedis = lmer(Firmicutesfam.incertaesedis.new ~ sex + age + pretreatment + treated_with_donor + clinical_outcome_wk14 * ns(timepoint.new, knots = 7) + (1 | subject_id), data = df)

#lme.slope.Firmicutesfam.incertaesedis = lmer(Firmicutesfam.incertaesedis.new ~ sex + age + pretreatment + treated_with_donor + clinical_outcome_wk14 * ns(timepoint.new, knots = 7) + (ns(timepoint.new, knots = 7) | subject_id), data = df)

glmm.int.Firmicutesfam.incertaesedis = glmmTMB(Firmicutesfam.incertaesedis.new ~ sex + age + pretreatment + treated_with_donor + clinical_outcome_wk14 * ns(timepoint.new, knots = 7) + (1 | subject_id), 
               family = gaussian, 
               data = df, 
               zi = ~ 1,
               REML = TRUE)

#glmm.slope.Firmicutesfam.incertaesedis = glmmTMB(Firmicutesfam.incertaesedis.new ~ sex + age + pretreatment + treated_with_donor + clinical_outcome_wk14 * ns(timepoint.new, knots = 7) + (ns(timepoint.new, knots = 7) | subject_id), 
               #family = gaussian, 
               #data = df, 
               #zi = ~ clinical_outcome_wk14 + ( 1 | subject_id),
               #REML = TRUE)

AICc(lme.int.Firmicutesfam.incertaesedis) # lower
AICc(glmm.int.Firmicutesfam.incertaesedis)
```

```{r}
# diagnostics lme
lme.Firmicutesfam.incertaesedis.diag = DHARMa::simulateResiduals(lme.int.Firmicutesfam.incertaesedis)

plot(lme.Firmicutesfam.incertaesedis.diag)

plotResiduals(lme.Firmicutesfam.incertaesedis.diag, form = df$clinical_outcome_wk14)
plotResiduals(lme.Firmicutesfam.incertaesedis.diag, form = df$timepoint.new)
plotResiduals(lme.Firmicutesfam.incertaesedis.diag, form = df$treated_with_donor)
plotResiduals(lme.Firmicutesfam.incertaesedis.diag, form = df$sex)
plotResiduals(lme.Firmicutesfam.incertaesedis.diag, form = df$age)
plotResiduals(lme.Firmicutesfam.incertaesedis.diag, form = df$pretreatment)

testZeroInflation(lme.Firmicutesfam.incertaesedis.diag)
testDispersion(lme.Firmicutesfam.incertaesedis.diag)
output1 = recalculateResiduals(lme.Firmicutesfam.incertaesedis.diag, group=df$timepoint.new)
testTemporalAutocorrelation(output1, time = unique(df$timepoint.new))
```

```{r}
# diagnostics zigmm
glmm.Firmicutesfam.incertaesedis.diag = DHARMa::simulateResiduals(glmm.int.Firmicutesfam.incertaesedis)

plot(glmm.Firmicutesfam.incertaesedis.diag)

plotResiduals(glmm.Firmicutesfam.incertaesedis.diag, form = df$clinical_outcome_wk14)
plotResiduals(glmm.Firmicutesfam.incertaesedis.diag, form = df$timepoint.new)
plotResiduals(glmm.Firmicutesfam.incertaesedis.diag, form = df$treated_with_donor)
plotResiduals(glmm.Firmicutesfam.incertaesedis.diag, form = df$sex)
plotResiduals(glmm.Firmicutesfam.incertaesedis.diag, form = df$age)
plotResiduals(glmm.Firmicutesfam.incertaesedis.diag, form = df$pretreatment)

testZeroInflation(glmm.Firmicutesfam.incertaesedis.diag)
testDispersion(glmm.Firmicutesfam.incertaesedis.diag)
output1 = recalculateResiduals(glmm.Firmicutesfam.incertaesedis.diag, group=df$timepoint.new)
testTemporalAutocorrelation(output1, time = unique(df$timepoint.new))
```
```{r}
wald.test(Sigma = vcov(lme.int.Firmicutesfam.incertaesedis), b = fixef(lme.int.Firmicutesfam.incertaesedis), Terms = c(6, 9, 10))$result

wald.test(Sigma = vcov(lme.int.Firmicutesfam.incertaesedis), b = fixef(lme.int.Firmicutesfam.incertaesedis), Terms = 2)$result # sex

wald.test(Sigma = vcov(lme.int.Firmicutesfam.incertaesedis), b = fixef(lme.int.Firmicutesfam.incertaesedis), Terms = 3)$result # age

wald.test(Sigma = vcov(lme.int.Firmicutesfam.incertaesedis), b = fixef(lme.int.Firmicutesfam.incertaesedis), Terms = 4)$result # pretreatment

wald.test(Sigma = vcov(lme.int.Firmicutesfam.incertaesedis), b = fixef(lme.int.Firmicutesfam.incertaesedis), Terms = 5)$result # donor
```

# Lachnospiraceae

## Transform the data 

```{r Histogramm Lachnospiraceae, echo=TRUE}
hist( df$Lachnospiraceae )
df$Lachnospiraceae.new = asin( sqrt( df$Lachnospiraceae ))
hist( df$Lachnospiraceae.new )
```

```{r Proportion Lachnospiraceae, echo=TRUE}
sum( df$Lachnospiraceae.new == 0 ) / length( df$Lachnospiraceae.new ) # proportion of zeros
```

```{r Plot data Lachnospiraceae, echo=TRUE}
# None responders
df.none = subset( df, clinical_outcome_wk14 == "None" ) # changed filter to subset

# Good responders
df.good = subset( df, clinical_outcome_wk14 == "Good" )
```

## Plot Lachnospiraceae
 
```{r Plot Lachnospiraceae, echo=TRUE, fig.width = 9}
a = xyplot( Lachnospiraceae.new ~ timepoint.new | clinical_outcome_wk14, data = df.none, xlab = "Timepoint", ylab = "Relative abundance (transformed)", main="Lachnospiraceae", 
           panel = function( x, y ) {
             panel.average( x, y, horizontal = FALSE, col = "#EC7063", lwd = 4 )
           },
           key = list( space = "right", 
                      lines = list( col = c( "#EC7063", "#82E0AA" ), lty = c( 1, 1 ), lwd = 2 ), 
                      text = list( c( "Non-Responders", "Responders" ))))

b = xyplot( Lachnospiraceae.new ~ timepoint.new | clinical_outcome_wk14, data = df.good, xlab = "Timepoint",
           panel = function( x, y ) {
             panel.average( x, y, horizontal = FALSE, col = "#82E0AA", lwd = 4, type = "l", lty = 1 )
           })

c = xyplot( Lachnospiraceae.new ~ timepoint.new | clinical_outcome_wk14, data = df.none, type = "p", col = "#EC7063" )

d = xyplot( Lachnospiraceae.new ~ timepoint.new | clinical_outcome_wk14, data = df.good, type = "p", col = "#82E0AA" )

Lachnospiraceae.plot = a + as.layer( b ) + as.layer( c ) + as.layer( d )
Lachnospiraceae.plot
```

## Choose between the lmer and glmmTMB models Lachnospiraceae

We have used time as a numeric value and added a spline at knots = 7.

```{r Mixed model Lachnospiraceae, echo=TRUE}
lme.int.Lachnospiraceae = lmer( Lachnospiraceae.new~ sex + age + pretreatment + treated_with_donor + clinical_outcome_wk14 * ns( timepoint.new.num, knots = 7 ) + ( 1 | subject_id ), data = df )

# lme.slope.Lachnospiraceaee = lmer( Lachnospiraceae.new ~ sex + age + pretreatment + treated_with_donor + clinical_outcome_wk14 * ns( timepoint.new.num, knots = 7 ) + ( ns( timepoint.new.num, knots = 7 ) | subject_id ),data = df )

glmm.int.Lachnospiraceae = glmmTMB( Lachnospiraceae.new ~ sex + age + pretreatment + treated_with_donor + clinical_outcome_wk14 * ns( timepoint.new.num, knots = 7 ) + ( 1 | subject_id ), 
               family = gaussian, 
               data = df, 
               zi = ~ 1,
               REML = TRUE )

# glmm.slope.Lachnospiraceae = glmmTMB( Lachnospiraceae.new ~ sex + age + pretreatment + treated_with_donor + clinical_outcome_wk14 * ns( timepoint.new.num, knots = 7 ) + ( ns( timepoint.new.num, knots = 7 ) | subject_id ), 
               #family = gaussian, 
               #data = df, 
               #zi = ~ 1,
               #REML = TRUE )

AICc( lme.int.Lachnospiraceae ) # lower
AICc( glmm.int.Lachnospiraceae )
```

# Diagnostics Lachnospiraceae

```{r Diagnostics Lachnospiraceae, eval=FALSE}
lme.Lachnospiraceae.diag = DHARMa::simulateResiduals( lme.int.Lachnospiraceae )
plot( lme.Lachnospiraceae.diag )

plotResiduals( lme.Lachnospiraceae.diag, form = df$clinical_outcome_wk14 )
plotResiduals( lme.Lachnospiraceae.diag, form = df$timepoint.new )
plotResiduals( lme.Lachnospiraceae.diag, form = df$treated_with_donor )
plotResiduals( lme.Lachnospiraceae.diag, form = df$sex ) 
plotResiduals( lme.Lachnospiraceae.diag, form = df$age )
plotResiduals( lme.Lachnospiraceae.diag, form = df$pretreatment )

testZeroInflation( lme.Lachnospiraceae.diag )
testDispersion( lme.Lachnospiraceae.diag )

output1 = recalculateResiduals( lme.Lachnospiraceae.diag, group = df$timepoint.new )
testTemporalAutocorrelation( output1, time = unique( df$timepoint.new ))
```

LMM was chosen due to its lower AICc value compared to ZIGMM as well as its adequate diagnostics.

```{r Results Lachnospiraceae}
plot_summs( lme.int.Lachnospiraceae )

wald.test( Sigma = vcov( lme.int.Lachnospiraceae ), b = fixef( lme.int.Lachnospiraceae ), Terms = c( 6, 9, 10 ))$result # clinical outcome wk14
wald.test( Sigma = vcov( lme.int.Lachnospiraceae ), b = fixef( lme.int.Lachnospiraceae ), Terms = 2 )$result # sex
wald.test( Sigma = vcov( lme.int.Lachnospiraceae ), b = fixef( lme.int.Lachnospiraceae ), Terms = 3 )$result # age
wald.test( Sigma = vcov( lme.int.Lachnospiraceae ), b = fixef( lme.int.Lachnospiraceae ), Terms = 4 )$result # pretreatment
wald.test( Sigma = vcov( lme.int.Lachnospiraceae ), b = fixef( lme.int.Lachnospiraceae ), Terms = 5 )$result # donor
```

# Oscillospiraceae

## Transform the data 

```{r Histogramm Oscillospiraceae, echo=TRUE}
hist( df$Oscillospiraceae )
df$Oscillospiraceae.new = asin( sqrt( df$Oscillospiraceae ))
hist( df$Oscillospiraceae.new )
```

```{r Proportion Oscillospiraceae, echo=TRUE}
sum( df$Oscillospiraceae.new == 0 ) / length( df$Oscillospiraceae.new ) # proportion of zeros
```

```{r Plot data Oscillospiraceae, echo=TRUE}
# None responders
df.none = subset( df, clinical_outcome_wk14 == "None" ) # changed filter to subset

# Good responders
df.good = subset( df, clinical_outcome_wk14 == "Good" )
```

## Plot Oscillospiraceae
 
```{r Plot Oscillospiraceae, echo=TRUE, fig.width = 9}
a = xyplot( Oscillospiraceae.new ~ timepoint.new | clinical_outcome_wk14, data = df.none, xlab = "Timepoint", ylab = "Relative abundance (transformed)", main="Oscillospiraceae", 
           panel = function( x, y ) {
             panel.average( x, y, horizontal = FALSE, col = "#EC7063", lwd = 4 )
           },
           key = list( space = "right", 
                      lines = list( col = c( "#EC7063", "#82E0AA" ), lty = c( 1, 1 ), lwd = 2 ), 
                      text = list( c( "Non-Responders", "Responders" ))))

b = xyplot( Oscillospiraceae.new ~ timepoint.new | clinical_outcome_wk14, data = df.good, xlab = "Timepoint",
           panel = function( x, y ) {
             panel.average( x, y, horizontal = FALSE, col = "#82E0AA", lwd = 4, type = "l", lty = 1 )
           })

c = xyplot( Oscillospiraceae.new ~ timepoint.new | clinical_outcome_wk14, data = df.none, type = "p", col = "#EC7063" )

d = xyplot( Oscillospiraceae.new ~ timepoint.new | clinical_outcome_wk14, data = df.good, type = "p", col = "#82E0AA" )

Oscillospiraceae.plot = a + as.layer( b ) + as.layer( c ) + as.layer( d )
Oscillospiraceae.plot
```

## Choose between the lmer and glmmTMB models Oscillospiraceae

We have used time as a numeric value and added a spline at knots = 7.

```{r Mixed model Oscillospiraceae, echo=TRUE}
lme.int.Oscillospiraceae = lme( Oscillospiraceae.new ~ sex + age + pretreatment + treated_with_donor + clinical_outcome_wk14 * ns( timepoint.new.num, knots = 7), 
               data = df, 
               random = ~ 1 | subject_id )

# lme.slope.Oscillospiraceae = lme( Oscillospiraceae.new ~ sex + age + pretreatment + treated_with_donor + clinical_outcome_wk14 * ns( timepoint.new.num, knots = 7), 
               #data = df, 
               #random = ~ ns( timepoint.new.num, knots = 7 ) | subject_id )

glmm.int.Oscillospiraceae = glmmTMB( Oscillospiraceae.new ~ sex + age + pretreatment + treated_with_donor + clinical_outcome_wk14 + ns( timepoint.new.num, knots = 7 ) + clinical_outcome_wk14 * ns( timepoint.new.num, knots = 7 ) + ( 1 | subject_id ), 
               family = gaussian, 
               data = df, 
               zi= ~ 1 )

#glmm.slope.Oscillospiraceae. = glmmTMB( Oscillospiraceae.new ~ sex + age + pretreatment + treated_with_donor + clinical_outcome_wk14 + ns( timepoint.new.num, knots = 7 ) + clinical_outcome_wk14 * ns( timepoint.new.num, knots = 7 ) + ( ns( timepoint.new.num, knots = 7 ) | subject_id ), 
               #family = gaussian, 
               #data = df, 
               #zi= ~ 1 )

AICc( lme.int.Oscillospiraceae )
AICc( glmm.int.Oscillospiraceae ) # lower
```

# Diagnostics Oscillospiraceae

```{r Diagnostics Oscillospiraceae, eval=FALSE}
lmer.int.Oscillospiraceae = lmer(Oscillospiraceae.new ~ sex + age + pretreatment + treated_with_donor + clinical_outcome_wk14 * ns(timepoint.new.num, knots = 7) + (1 | subject_id), data = df)

lme.Oscillospiraceae.diag = DHARMa::simulateResiduals( lmer.int.Oscillospiraceae )
plot( lme.Oscillospiraceae.diag )

plotResiduals( lme.Oscillospiraceae.diag, form = df$clinical_outcome_wk14 )
plotResiduals( lme.Oscillospiraceae.diag, form = df$timepoint.new )
plotResiduals( lme.Oscillospiraceae.diag, form = df$treated_with_donor )
plotResiduals( lme.Oscillospiraceae.diag, form = df$sex ) 
plotResiduals( lme.Oscillospiraceae.diag, form = df$age )
plotResiduals( lme.Oscillospiraceae.diag, form = df$pretreatment )

testZeroInflation( lme.Oscillospiraceae.diag )
testDispersion( lme.Oscillospiraceae.diag )

output1 = recalculateResiduals( lme.Oscillospiraceae.diag, group = df$timepoint.new )
testTemporalAutocorrelation( output1, time = unique( df$timepoint.new ))
```

LMM was chosen due to better diagnostics than ZIGMM despite ZIGMM having a lower AICc value. 

```{r Results Oscillospiraceae}
plot_summs( lme.int.Oscillospiraceae )

wald.test( Sigma = vcov( lme.int.Oscillospiraceae ), b = fixef( lme.int.Oscillospiraceae ), Terms = c( 6, 9, 10 ))$result # clinical outcome wk14
wald.test( Sigma = vcov( lme.int.Oscillospiraceae ), b = fixef( lme.int.Oscillospiraceae ), Terms = 2 )$result # sex
wald.test( Sigma = vcov( lme.int.Oscillospiraceae ), b = fixef( lme.int.Oscillospiraceae ), Terms = 3 )$result # age
wald.test( Sigma = vcov( lme.int.Oscillospiraceae ), b = fixef( lme.int.Oscillospiraceae ), Terms = 4 )$result # pretreatment
wald.test( Sigma = vcov( lme.int.Oscillospiraceae ), b = fixef( lme.int.Oscillospiraceae ), Terms = 5 )$result # donor
```

# Prevotellaceae

## Transform the data 

```{r Histogramm Prevotellaceae, echo=TRUE}
hist( df$Prevotellaceae )
df$Prevotellaceae.new = asin( sqrt( df$Prevotellaceae ))
hist( df$Prevotellaceae.new )
```

```{r Proportion Prevotellaceae, echo=TRUE}
sum( df$Prevotellaceae.new == 0 ) / length( df$Prevotellaceae.new ) # proportion of zeros
```

```{r Plot data Prevotellaceae, echo=TRUE}
# None responders
df.none = subset( df, clinical_outcome_wk14 == "None" ) # changed filter to subset

# Good responders
df.good = subset( df, clinical_outcome_wk14 == "Good" )
```

## Plot Prevotellaceae

```{r Plot Prevotellaceae, echo=TRUE, fig.width = 9}
a = xyplot( Prevotellaceae.new ~ timepoint.new | clinical_outcome_wk14, data = df.none, xlab = "Timepoint", ylab = "Relative abundance (transformed)", main="Prevotellaceae", 
           panel = function( x, y ) {
             panel.average( x, y, horizontal = FALSE, col = "#EC7063", lwd = 4 )
           },
           key = list( space = "right", 
                      lines = list( col = c( "#EC7063", "#82E0AA" ), lty = c( 1, 1 ), lwd = 2 ), 
                      text = list( c( "Non-Responders", "Responders" ))))

b = xyplot( Prevotellaceae.new ~ timepoint.new | clinical_outcome_wk14, data = df.good, xlab = "Timepoint",
           panel = function( x, y ) {
             panel.average( x, y, horizontal = FALSE, col = "#82E0AA", lwd = 4, type = "l", lty = 1 )
           })

c = xyplot( Prevotellaceae.new ~ timepoint.new | clinical_outcome_wk14, data = df.none, type = "p", col = "#EC7063" )

d = xyplot( Prevotellaceae.new ~ timepoint.new | clinical_outcome_wk14, data = df.good, type = "p", col = "#82E0AA" )

Prevotellaceae.plot = a + as.layer( b ) + as.layer( c ) + as.layer( d )
Prevotellaceae.plot
```

## Choose between the lmer and glmmTMB models Prevotellaceae

We have used time as a numeric value and added a spline at knots = 7.

```{r Mixed model Prevotellaceae, echo=TRUE}
lme.int.Prevotellaceae = lmer( Prevotellaceae.new ~ sex + age + pretreatment + treated_with_donor + clinical_outcome_wk14 * ns( timepoint.new.num, knots = 7 ) + ( 1 | subject_id ), data = df )

# lme.slope.Prevotellaceae = lmer( Prevotellaceae.new ~ sex + age + pretreatment + treated_with_donor + clinical_outcome_wk14 * ns(timepoint.new.num, knots = 7) + ( ns( timepoint.new.num, knots = 7 ) | subject_id ), data = df )

glmm.int.Prevotellaceae = glmmTMB( Prevotellaceae.new ~ sex + age + pretreatment + treated_with_donor + clinical_outcome_wk14 * ns( timepoint.new.num, knots = 7 ) + ( 1 | subject_id ), 
               family = gaussian, 
               data = df, 
               zi = ~ 1,
               REML = TRUE )

# glmm.slope.Prevotellaceae = glmmTMB( Prevotellaceae.new ~ sex + age + pretreatment + treated_with_donor + clinical_outcome_wk14 * ns( timepoint.new.num, knots = 7 ) + ( ns( timepoint.new.num, knots = 7 ) | subject_id ), 
               #family = gaussian, 
               #data = df, 
               #zi = ~ 1,
               #REML = TRUE )

AICc( lme.int.Prevotellaceae ) # lower
AICc( glmm.int.Prevotellaceae )
```

# Diagnostics Prevotellaceae

```{r Diagnostics Prevotellaceae, eval=FALSE}
lme.Prevotellaceae.diag = DHARMa::simulateResiduals( lme.int.Prevotellaceae )
plot( lme.Prevotellaceae.diag )

plotResiduals( lme.Prevotellaceae.diag, form = df$clinical_outcome_wk14 )
plotResiduals( lme.Prevotellaceae.diag, form = df$timepoint.new )
plotResiduals( lme.Prevotellaceae.diag, form = df$treated_with_donor )
plotResiduals( lme.Prevotellaceae.diag, form = df$sex ) 
plotResiduals( lme.Prevotellaceae.diag, form = df$age )
plotResiduals( lme.Prevotellaceae.diag, form = df$pretreatment )

testZeroInflation( lme.Prevotellaceae.diag )
testDispersion( lme.Prevotellaceae.diag )

output1 = recalculateResiduals( lme.Prevotellaceae.diag, group = df$timepoint.new )
testTemporalAutocorrelation( output1, time = unique( df$timepoint.new ))
```

As there were no major differences in diagnostics between LMM and ZIGMM, the LMM model was chosen due to its lower AICc value. 

```{r Results Prevotellaceae}
plot_summs( lme.int.Prevotellaceae )

wald.test( Sigma = vcov( lme.int.Prevotellaceae ), b = fixef( lme.int.Prevotellaceae ), Terms = c( 6, 9, 10 ))$result # clinical outcome wk14
wald.test( Sigma = vcov( lme.int.Prevotellaceae ), b = fixef( lme.int.Prevotellaceae ), Terms = 2 )$result # sex
wald.test( Sigma = vcov( lme.int.Prevotellaceae ), b = fixef( lme.int.Prevotellaceae ), Terms = 3 )$result # age
wald.test( Sigma = vcov( lme.int.Prevotellaceae ), b = fixef( lme.int.Prevotellaceae ), Terms = 4 )$result # pretreatment
wald.test( Sigma = vcov( lme.int.Prevotellaceae ), b = fixef( lme.int.Prevotellaceae ), Terms = 5 )$result # donor
```

# Rikenellaceae

## Transform the data 

```{r Histogramm Rikenellaceae, echo=TRUE}
hist( df$Rikenellaceae )
df$Rikenellaceae.new = asin( sqrt( df$Rikenellaceae ))
hist( df$Rikenellaceae.new )
```

```{r Proportion Rikenellaceae, echo=TRUE}
sum( df$Rikenellaceae.new == 0 ) / length( df$Rikenellaceae.new ) # proportion of zeros
```

```{r Plot data Rikenellaceae, echo=TRUE}
# None responders
df.none = subset( df, clinical_outcome_wk14 == "None" ) # changed filter to subset

# Good responders
df.good = subset( df, clinical_outcome_wk14 == "Good" )
```

## Plot Rikenellaceae
 
```{r Plot Rikenellaceae, echo=TRUE, fig.width = 9}
a = xyplot( Rikenellaceae.new ~ timepoint.new | clinical_outcome_wk14, data = df.none, xlab = "Timepoint", ylab = "Relative abundance (transformed)", main="Rikenellaceae", 
           panel = function( x, y ) {
             panel.average( x, y, horizontal = FALSE, col = "#EC7063", lwd = 4 )
           },
           key = list( space = "right", 
                      lines = list( col = c( "#EC7063", "#82E0AA" ), lty = c( 1, 1 ), lwd = 2 ), 
                      text = list( c( "Non-Responders", "Responders" ))))

b = xyplot( Rikenellaceae.new ~ timepoint.new | clinical_outcome_wk14, data = df.good, xlab = "Timepoint",
           panel = function( x, y ) {
             panel.average( x, y, horizontal = FALSE, col = "#82E0AA", lwd = 4, type = "l", lty = 1 )
           })

c = xyplot( Rikenellaceae.new ~ timepoint.new | clinical_outcome_wk14, data = df.none, type = "p", col = "#EC7063" )

d = xyplot( Rikenellaceae.new ~ timepoint.new | clinical_outcome_wk14, data = df.good, type = "p", col = "#82E0AA" )

Rikenellaceae.plot = a + as.layer( b ) + as.layer( c ) + as.layer( d )
Rikenellaceae.plot
```

## Choose between the lmer and glmmTMB models Rikenellaceae

We have used time as a numeric value and added a spline at knots = 7.

```{r Mixed model Rikenellaceae, echo=TRUE}
lme.int.Rikenellaceae = lme( Rikenellaceae.new ~ sex + age + pretreatment + treated_with_donor + clinical_outcome_wk14 * ns( timepoint.new.num, knots = 7 ), 
               data = df, 
               random = ~ 1 | subject_id )

# lme.slope.Rikenellaceae = lme( Rikenellaceae.new ~ sex + age + pretreatment + treated_with_donor + clinical_outcome_wk14 * ns( timepoint.new.num, knots = 7 ), 
               #data = df, 
               #random = ~ ns( timepoint.new.num, knots = 7 ) | subject_id )

glmm.int.Rikenellaceae = glmmTMB( Rikenellaceae.new ~ sex + age + pretreatment + treated_with_donor + clinical_outcome_wk14 + ns( timepoint.new.num, knots = 7 ) + clinical_outcome_wk14 * ns( timepoint.new.num, knots = 7 ) + ( 1 | subject_id ), 
               family = gaussian, 
               data = df, 
               zi= ~ clinical_outcome_wk14 + ns( timepoint.new.num, knots = 7 ) + ( 1 | subject_id ))

#glmm.slope.Rikenellaceae = glmmTMB(Rikenellaceae.new ~ sex + age + pretreatment + treated_with_donor + clinical_outcome_wk14 + ns(timepoint.new.num, knots = 7) + clinical_outcome_wk14 * ns(timepoint.new.num, knots = 7) + (ns(timepoint.new.num, knots = 7) | subject_id), 
               #family = gaussian, 
               #data = df, 
               #zi= ~ clinical_outcome_wk14 + ns(timepoint.new.num, knots = 7) + ( 1 | subject_id))

AICc( lme.int.Rikenellaceae ) 
AICc( glmm.int.Rikenellaceae ) # lower
```

# Diagnostics Rikenellaceae

```{r Diagnostics Rikenellaceae, eval=FALSE}
glmm.Rikenellaceae.diag = DHARMa::simulateResiduals( glmm.int.Rikenellaceae )
plot( glmm.Rikenellaceae.diag )

plotResiduals( glmm.Rikenellaceae.diag, form = df$clinical_outcome_wk14 )
plotResiduals( glmm.Rikenellaceae.diag, form = df$timepoint.new )
plotResiduals( glmm.Rikenellaceae.diag, form = df$treated_with_donor )
plotResiduals( glmm.Rikenellaceae.diag, form = df$sex ) 
plotResiduals( glmm.Rikenellaceae.diag, form = df$age )
plotResiduals( glmm.Rikenellaceae.diag, form = df$pretreatment )

testZeroInflation( glmm.Rikenellaceae.diag )
testDispersion( glmm.Rikenellaceae.diag )

output1 = recalculateResiduals( glmm.Rikenellaceae.diag, group = df$timepoint.new )
testTemporalAutocorrelation( output1, time = unique( df$timepoint.new ))
```

ZIGMM was chosen due to a lower AICc value comapred to LMM. However, there are problems with diagnostics with both models.  

```{r Results Rikenellaceae, fig.width = 9}
conf.int = data.frame( confint( glmm.int.Rikenellaceae ))
colnames( conf.int )[ 1 ] = "conf.low"
colnames( conf.int )[ 2 ] = "conf.high"
conf.int = conf.int[ -c( 11:21 ), ]

clostridfam = broom.mixed::tidy( glmm.int.Rikenellaceae )[ , -c( 1 )]
clostridfam = clostridfam[ -c( 11:23 ), ]
clostridfam = clostridfam[ , -c( 1 )]
effect = "fixed"

clostridfam = data.frame( model = "Clostridiales incertaesedis", effect = effect, clostridfam, conf.low = conf.int$conf.low, conf.high = conf.int$conf.high )
clostridfam = clostridfam[ , -c( 8 )]
clostridfam = as_tibble( clostridfam )

dwplot( clostridfam ) + xlab( "Coefficient Estimate" ) + geom_vline( xintercept = 0, colour = "grey60", linetype = 2 ) + theme_bw() + theme( legend.title = element_blank())

wald.test( Sigma = vcov( glmm.int.Rikenellaceae )[[ "cond" ]], b = fixef( glmm.int.Rikenellaceae )[[ "cond" ]], Terms = c( 6, 9, 10 ))$result

wald.test( Sigma = vcov( glmm.int.Rikenellaceae )[[ "cond" ]], b = fixef( glmm.int.Rikenellaceae )[[ "cond" ]], Terms = 2 )$result # sex

wald.test( Sigma = vcov( glmm.int.Rikenellaceae )[[ "cond" ]], b = fixef( glmm.int.Rikenellaceae )[[ "cond" ]], Terms = 3 )$result # age

wald.test( Sigma = vcov( glmm.int.Rikenellaceae )[[ "cond" ]], b = fixef( glmm.int.Rikenellaceae )[[ "cond" ]], Terms = 4 )$result # pretreatment

wald.test( Sigma = vcov( glmm.int.Rikenellaceae )[[ "cond" ]], b = fixef( glmm.int.Rikenellaceae )[[ "cond" ]], Terms = 5 )$result # donor
```

# Ruminococcaceae

## Transform the data 

```{r Histogramm Ruminococcaceae, echo=TRUE}
hist( df$Ruminococcaceae )
# No transformation needed
```

```{r Proportion Ruminococcaceae, echo=TRUE}
sum( df$Ruminococcaceae == 0 ) / length( df$Ruminococcaceae ) # proportion of zeros
```

```{r Plot data Ruminococcaceae, echo=TRUE}
# None responders
df.none = subset( df, clinical_outcome_wk14 == "None" ) # changed filter to subset

# Good responders
df.good = subset( df, clinical_outcome_wk14 == "Good" )
```

## Plot Ruminococcaceae
 
```{r Plot Ruminococcaceae, echo=TRUE, fig.width = 9}
a = xyplot( Ruminococcaceae ~ timepoint.new | clinical_outcome_wk14, data = df.none, xlab = "Timepoint", ylab = "Relative abundance (transformed)", main="Ruminococcaceae", 
           panel = function( x, y ) {
             panel.average( x, y, horizontal = FALSE, col = "#EC7063", lwd = 4 )
           },
           key = list( space = "right", 
                      lines = list( col = c( "#EC7063", "#82E0AA" ), lty = c( 1, 1 ), lwd = 2 ), 
                      text = list( c( "Non-Responders", "Responders" ))))

b = xyplot( Ruminococcaceae ~ timepoint.new | clinical_outcome_wk14, data = df.good, xlab = "Timepoint",
           panel = function( x, y ) {
             panel.average( x, y, horizontal = FALSE, col = "#82E0AA", lwd = 4, type = "l", lty = 1 )
           })

c = xyplot( Ruminococcaceae ~ timepoint.new | clinical_outcome_wk14, data = df.none, type = "p", col = "#EC7063" )

d = xyplot( Ruminococcaceae ~ timepoint.new | clinical_outcome_wk14, data = df.good, type = "p", col = "#82E0AA" )

Ruminococcaceae.plot = a + as.layer( b ) + as.layer( c ) + as.layer( d )
Ruminococcaceae.plot
```

## Choose between the lmer and glmmTMB models Ruminococcaceae

We have used time as a numeric value and added a spline at knots = 7.

```{r Mixed model Ruminococcaceae, echo=TRUE}
lme.int.Ruminococcaceae = lme( Ruminococcaceae ~ sex + age + pretreatment + treated_with_donor + clinical_outcome_wk14 * ns( timepoint.new.num, knots = 7 ), 
               data = df, 
               random = ~ 1 | subject_id )

# lme.slope.Ruminococcaceae = lme( Ruminococcaceae ~ sex + age + pretreatment + treated_with_donor + clinical_outcome_wk14 * ns( timepoint.new.num, knots = 7 ), 
               #data = df, 
               #random = ~ ns( timepoint.new.num, knots = 7 ) | subject_id )

summary( lme.int.Ruminococcaceae )
```

# Diagnostics Ruminococcaceae

```{r Diagnostics Ruminococcaceae, eval=FALSE}
lmer.int.Ruminococcaceae = lmer( Ruminococcaceae ~ sex + age + pretreatment + treated_with_donor + clinical_outcome_wk14 * ns( timepoint.new.num, knots = 7 ) + ( 1 | subject_id ),
               data = df )

lmer.Ruminococcaceae.diag = DHARMa::simulateResiduals( lmer.int.Ruminococcaceae )
plot( lmer.Ruminococcaceae.diag )

plotResiduals( lmer.Ruminococcaceae.diag, form = df$clinical_outcome_wk14 )
plotResiduals( lmer.Ruminococcaceae.diag, form = df$timepoint.new )
plotResiduals( lmer.Ruminococcaceae.diag, form = df$treated_with_donor )
plotResiduals( lmer.Ruminococcaceae.diag, form = df$sex ) 
plotResiduals( lmer.Ruminococcaceae.diag, form = df$age )
plotResiduals( lmer.Ruminococcaceae.diag, form = df$pretreatment )

testZeroInflation( lmer.Ruminococcaceae.diag )
testDispersion( lmer.Ruminococcaceae.diag )

output1 = recalculateResiduals( lmer.Ruminococcaceae.diag, group = df$timepoint.new )
testTemporalAutocorrelation( output1, time = unique( df$timepoint.new ))
```

LMM was chosen due to its lower AICc value and adequate diagnostics. 

```{r Results Ruminococcaceae}
plot_summs( lme.int.Ruminococcaceae )

wald.test( Sigma = vcov( lme.int.Ruminococcaceae ), b = fixef( lme.int.Ruminococcaceae ), Terms = c( 6, 9, 10 ))$result # clinical outcome wk14
wald.test( Sigma = vcov( lme.int.Ruminococcaceae ), b = fixef( lme.int.Ruminococcaceae ), Terms = 2 )$result # sex
wald.test( Sigma = vcov( lme.int.Ruminococcaceae ), b = fixef( lme.int.Ruminococcaceae ), Terms = 3 )$result # age
wald.test( Sigma = vcov( lme.int.Ruminococcaceae ), b = fixef( lme.int.Ruminococcaceae ), Terms = 4 )$result # pretreatment
wald.test( Sigma = vcov( lme.int.Ruminococcaceae ), b = fixef( lme.int.Ruminococcaceae ), Terms = 5 )$result # donor
```


# Sutterellaceae

```{r}
sum(df$Sutterellaceae == 0) / length(df$Sutterellaceae) 
```

```{r}
xyplot(Sutterellaceae ~ timepoint.new | clinical_outcome_wk14, data = df, xlab = "Timepoint", ylab = "Sutterellaceae", ylim = c(0, 0.05),
           panel = function(x, y) {
            panel.average(x, y, horizontal = FALSE, col = "plum", lwd = 4)
       })

hist(df$Sutterellaceae)

df$Sutterellaceae.new = asin(sqrt(df$Sutterellaceae))
hist(df$Sutterellaceae.new)

xyplot(Sutterellaceae.new ~ timepoint.new | clinical_outcome_wk14, data = df, xlab = "Timepoint", ylab = "Sutterellaceae", ylim = c(0, 0.2),
           panel = function(x, y) {
            panel.average(x, y, horizontal = FALSE, col = "plum", lwd = 4)
       })
```

```{r}
lme.int.Sutterellaceae = lmer(Sutterellaceae.new ~ sex + age + pretreatment + treated_with_donor + clinical_outcome_wk14 * ns(timepoint.new, knots = 7) + (1 | subject_id), data = df)

#lme.slope.Sutterellaceae = lmer(Sutterellaceae.new ~ sex + age + pretreatment + treated_with_donor + clinical_outcome_wk14 * ns(timepoint.new, knots = 7) + (ns(timepoint.new, knots = 7) | subject_id), data = df)

glmm.int.Sutterellaceae = glmmTMB(Sutterellaceae.new ~ sex + age + pretreatment + treated_with_donor + clinical_outcome_wk14 * ns(timepoint.new, knots = 7) + (1 | subject_id), 
               family = gaussian, 
               data = df, 
               zi= ~ 1)

#glmm.slope.Sutterellaceae = glmmTMB(Sutterellaceae.new ~ sex + age + pretreatment + treated_with_donor + clinical_outcome_wk14 * ns(timepoint.new, knots = 7) + (ns(timepoint.new, knots = 7) | subject_id), 
               #family = gaussian, 
               #data = df, 
               #zi= ~ 1)

AICc(lme.int.Sutterellaceae)
AICc(glmm.int.Sutterellaceae) # lower
```

```{r}
glmm.Sutterellaceae.diag = DHARMa::simulateResiduals(glmm.int.Sutterellaceae)
plot(glmm.Sutterellaceae.diag)

plotResiduals(glmm.Sutterellaceae.diag, form = df$clinical_outcome_wk14)
plotResiduals(glmm.Sutterellaceae.diag, form = df$timepoint.new)
plotResiduals(glmm.Sutterellaceae.diag, form = df$treated_with_donor)
plotResiduals(glmm.Sutterellaceae.diag, form = df$sex)
plotResiduals(glmm.Sutterellaceae.diag, form = df$age)
plotResiduals(glmm.Sutterellaceae.diag, form = df$pretreatment)

testZeroInflation(glmm.Sutterellaceae.diag)
testDispersion(glmm.Sutterellaceae.diag)
output1 = recalculateResiduals(glmm.Sutterellaceae.diag, group=df$timepoint.new)
testTemporalAutocorrelation(output1, time = unique(df$timepoint.new))
```

```{r}
lme.Sutterellaceae.diag = DHARMa::simulateResiduals(lme.int.Sutterellaceae)
plot(lme.Sutterellaceae.diag)

plotResiduals(lme.Sutterellaceae.diag, form = df$clinical_outcome_wk14)
plotResiduals(lme.Sutterellaceae.diag, form = df$timepoint.new)
plotResiduals(lme.Sutterellaceae.diag, form = df$treated_with_donor)
plotResiduals(lme.Sutterellaceae.diag, form = df$sex)
plotResiduals(lme.Sutterellaceae.diag, form = df$age)
plotResiduals(lme.Sutterellaceae.diag, form = df$pretreatment)

testZeroInflation(lme.Sutterellaceae.diag)
testDispersion(lme.Sutterellaceae.diag)
output1 = recalculateResiduals(lme.Sutterellaceae.diag, group=df$timepoint.new)
testTemporalAutocorrelation(output1, time = unique(df$timepoint.new))
```


```{r}
wald.test(Sigma = vcov(glmm.int.Sutterellaceae)[["cond"]], b = fixef(glmm.int.Sutterellaceae)[["cond"]], Terms = c(6, 9, 10))$result

wald.test(Sigma = vcov(glmm.int.Sutterellaceae)[["cond"]], b = fixef(glmm.int.Sutterellaceae)[["cond"]], Terms = 2)$result # sex

wald.test(Sigma = vcov(glmm.int.Sutterellaceae)[["cond"]], b = fixef(glmm.int.Sutterellaceae)[["cond"]], Terms = 3)$result # age

wald.test(Sigma = vcov(glmm.int.Sutterellaceae)[["cond"]], b = fixef(glmm.int.Sutterellaceae)[["cond"]], Terms = 4)$result # pretreatment

wald.test(Sigma = vcov(glmm.int.Sutterellaceae)[["cond"]], b = fixef(glmm.int.Sutterellaceae)[["cond"]], Terms = 5)$result # donor
```

# Veillonellaceae

## Transform the data 

```{r Histogramm Veillonellaceae, echo=TRUE}
hist( df$Veillonellaceae )
df$Veillonellaceae.new = asin( sqrt( df$Veillonellaceae ))
hist( df$Veillonellaceae.new )
```

```{r Proportion Veillonellaceae, echo=TRUE}
sum( df$Veillonellaceae.new == 0 ) / length( df$Veillonellaceae.new ) # proportion of zeros
```

```{r Plot data Veillonellaceae, echo=TRUE}
# None responders
df.none = subset( df, clinical_outcome_wk14 == "None" ) # changed filter to subset

# Good responders
df.good = subset( df, clinical_outcome_wk14 == "Good" )
```

## Plot Veillonellaceae
 
```{r Plot Veillonellaceae, echo=TRUE, fig.width = 9}
a = xyplot( Veillonellaceae.new ~ timepoint.new | clinical_outcome_wk14, data = df.none, xlab = "Timepoint", ylab = "Relative abundance (transformed)", main="Veillonellaceae", 
           panel = function( x, y ) {
             panel.average( x, y, horizontal = FALSE, col = "#EC7063", lwd = 4 )
           },
           key = list( space = "right", 
                      lines = list( col = c( "#EC7063", "#82E0AA" ), lty = c( 1, 1 ), lwd = 2 ), 
                      text = list( c( "Non-Responders", "Responders" ))))

b = xyplot( Veillonellaceae.new ~ timepoint.new | clinical_outcome_wk14, data = df.good, xlab = "Timepoint",
           panel = function( x, y ) {
             panel.average( x, y, horizontal = FALSE, col = "#82E0AA", lwd = 4, type = "l", lty = 1 )
           })

c = xyplot( Veillonellaceae.new ~ timepoint.new | clinical_outcome_wk14, data = df.none, type = "p", col = "#EC7063" )

d = xyplot( Veillonellaceae.new ~ timepoint.new | clinical_outcome_wk14, data = df.good, type = "p", col = "#82E0AA" )

Veillonellaceae.plot = a + as.layer( b ) + as.layer( c ) + as.layer( d )
Veillonellaceae.plot
```

## Choose between the lmer and glmmTMB models Veillonellaceae

We have used time as a numeric value and added a spline at knots = 7.

```{r Mixed model Veillonellaceae, echo=TRUE}
lme.int.Veillonellaceae = lmer( Veillonellaceae.new ~ sex + age + pretreatment + treated_with_donor + clinical_outcome_wk14 * ns( timepoint.new.num, knots = 7 ) + ( 1 | subject_id ), 
               data = df )

lme.slope.Veillonellaceae = lmer( Veillonellaceae.new ~ sex + age + pretreatment + treated_with_donor + clinical_outcome_wk14 * ns( timepoint.new.num, knots = 7 ) + ( ns( timepoint.new.num, knots = 7 ) | subject_id ), 
               data = df )

# test random slopes vs random intercepts
anova_results = anova( lme.int.Veillonellaceae, lme.slope.Veillonellaceae )
mixtureLRT = anova_results[[ "Chisq" ]][ 2 ]
0.5 * pchisq( mixtureLRT, 1, lower.tail = FALSE ) + 0.5 * pchisq( mixtureLRT, 2, lower.tail = FALSE )
# use random intercepts

glmm.int.Veillonellaceae = glmmTMB( Veillonellaceae.new ~ sex + age + pretreatment + treated_with_donor + clinical_outcome_wk14 * ns(timepoint.new.num, knots = 7 ) + ( 1 | subject_id ), 
               family = gaussian, 
               data = df, 
               zi= ~ 1 ) # can't make this more complicated

glmm.slope.Veillonellaceae = glmmTMB( Veillonellaceae.new ~ sex + age + pretreatment + treated_with_donor + clinical_outcome_wk14 * ns( timepoint.new.num, knots = 7 ) + ( ns( timepoint.new.num, knots = 7 ) | subject_id ), 
               family = gaussian, 
               data = df, 
               zi= ~ 1 ) 

# test random slopes vs random intercepts
anova_results = anova( glmm.int.Veillonellaceae, glmm.slope.Veillonellaceae )
mixtureLRT = anova_results[[ "Chisq" ]][ 2 ]
0.5 * pchisq( mixtureLRT, 1, lower.tail = FALSE ) + 0.5 * pchisq( mixtureLRT, 2, lower.tail = FALSE )
# use random slopes model

AICc( lme.int.Veillonellaceae )
AIC( lme.slope.Veillonellaceae )
AICc( glmm.int.Veillonellaceae )
AICc( glmm.slope.Veillonellaceae ) # lower
```

# Diagnostics Veillonellaceae

```{r Diagnostics Veillonellaceae, eval=FALSE}
glmm.Veillonellaceae.diag = DHARMa::simulateResiduals( glmm.slope.Veillonellaceae )
plot( glmm.Veillonellaceae.diag )

plotResiduals( glmm.Veillonellaceae.diag, form = df$clinical_outcome_wk14 )
plotResiduals( glmm.Veillonellaceae.diag, form = df$timepoint.new )
plotResiduals( glmm.Veillonellaceae.diag, form = df$treated_with_donor )
plotResiduals( glmm.Veillonellaceae.diag, form = df$sex ) 
plotResiduals( glmm.Veillonellaceae.diag, form = df$age )
plotResiduals( glmm.Veillonellaceae.diag, form = df$pretreatment )

testZeroInflation( glmm.Veillonellaceae.diag )
testDispersion( glmm.Veillonellaceae.diag )

output1 = recalculateResiduals( glmm.Veillonellaceae.diag, group = df$timepoint.new )
testTemporalAutocorrelation( output1, time = unique( df$timepoint.new ))
```

ZIGMM was chosen due to a lower AICc value. Diagnostics between LMM and ZIGMM were similar. 

```{r Results Veillonellaceae, fig.width = 9}
conf.int = data.frame( confint( glmm.slope.Veillonellaceae ))
colnames( conf.int )[ 1 ] = "conf.low"
colnames( conf.int )[ 2 ] = "conf.high"
conf.int = conf.int[ -c( 11:21 ), ]

clostridfam = broom.mixed::tidy( glmm.slope.Veillonellaceae )[ , -c( 1 )]
clostridfam = clostridfam[ -c( 11:23 ), ]
clostridfam = clostridfam[ , -c( 1 )]
effect = "fixed"

clostridfam = data.frame( model = "Clostridiales incertaesedis", effect = effect, clostridfam, conf.low = conf.int$conf.low, conf.high = conf.int$conf.high )
clostridfam = clostridfam[ , -c( 8 )]
clostridfam = as_tibble( clostridfam )

dwplot( clostridfam ) + xlab( "Coefficient Estimate" ) + geom_vline( xintercept = 0, colour = "grey60", linetype = 2 ) + theme_bw() + theme( legend.title = element_blank())

wald.test( Sigma = vcov( glmm.slope.Veillonellaceae )[[ "cond" ]], b = fixef( glmm.slope.Veillonellaceae )[[ "cond" ]], Terms = c( 6, 9, 10 ))$result

wald.test( Sigma = vcov( glmm.slope.Veillonellaceae )[[ "cond" ]], b = fixef( glmm.slope.Veillonellaceae )[[ "cond" ]], Terms = 2 )$result # sex

wald.test( Sigma = vcov( glmm.slope.Veillonellaceae )[[ "cond" ]], b = fixef( glmm.slope.Veillonellaceae )[[ "cond" ]], Terms = 3 )$result # age

wald.test( Sigma = vcov( glmm.slope.Veillonellaceae )[[ "cond" ]], b = fixef( glmm.slope.Veillonellaceae )[[ "cond" ]], Terms = 4 )$result # pretreatment

wald.test( Sigma = vcov( glmm.slope.Veillonellaceae )[[ "cond" ]], b = fixef( glmm.slope.Veillonellaceae )[[ "cond" ]], Terms = 5 )$result # donor
```
# Coefficients Plot - Comparison

```{r Coefficients plot data, echo=TRUE}
# Bacteroidaceae
bacteroid = broom.mixed::tidy( lme.int.Bacteroidaceae, conf.int = TRUE )
bacteroid = bacteroid[ -c( 11:12 ), ]
model = "Bacteroidaceae"
bacteroid = data.frame( model = model, bacteroid )
bacteroid = as_tibble( bacteroid )

# Bacteroidalesfam.incertaesedis

# Bifidobacteriaceae

# Clostridiaceae
conf.int = data.frame( confint( glmm.int.Clostridiaceae ))
colnames( conf.int )[ 1 ] = "conf.low"
colnames( conf.int )[ 2 ] = "conf.high"
conf.int = conf.int[ -c( 11:21 ), ]

clostrid = broom.mixed::tidy( glmm.int.Clostridiaceae )[ , -c( 1 )]
clostrid = clostrid[ -c( 11:23 ), ]
clostrid = clostrid[ , -c( 1 )]
effect = "fixed"

clostrid = data.frame( model = "Clostridiaceae", effect = effect, clostrid, conf.low = conf.int$conf.low, conf.high = conf.int$conf.high )
clostrid = clostrid[ , -c( 8 )]
clostrid = as_tibble( clostrid )

# Clostridiales incertaesedis
conf.int = data.frame( confint( glmm.slope.Clostridialesfam.incertaesedis ))
colnames( conf.int )[ 1 ] = "conf.low"
colnames( conf.int )[ 2 ] = "conf.high"
conf.int = conf.int[ -c( 11:21 ), ]

clostridfam = broom.mixed::tidy( glmm.slope.Clostridialesfam.incertaesedis )[ , -c( 1 )]
clostridfam = clostridfam[ -c( 11:23 ), ]
clostridfam = clostridfam[ , -c( 1 )]
effect = "fixed"

clostridfam = data.frame( model = "Clostridiales incertaesedis", effect = effect, clostridfam, conf.low = conf.int$conf.low, conf.high = conf.int$conf.high )
clostridfam = clostridfam[ , -c( 8 )]
clostridfam = as_tibble( clostridfam )

# Coriobacteriaceae

# Eubacteriaceae 

# Firmicutesfam.incertaesedis

# Lachnospiraceae
lachno = broom.mixed::tidy( lme.int.Lachnospiraceae, conf.int = TRUE )
lachno = lachno[ -c( 11:12 ), -c( 7:8 )]
model = "Lachnospiraceae"
lachno = data.frame( model = model, lachno )
lachno = as_tibble( lachno )

# Oscillospiraceae
oscillo = broom.mixed::tidy( lme.int.Oscillospiraceae, conf.int = TRUE )
oscillo = oscillo[ -c( 11:12 ), -c( 6,8 )]
model = "Oscillospiraceae"
oscillo = data.frame( model = model, oscillo )
oscillo = as_tibble( oscillo )

# Prevotellaceae
prevo = broom.mixed::tidy( lme.int.Prevotellaceae, conf.int = TRUE )
prevo = prevo[ -c( 11:12 ), -c( 7:8 )]
model = "Prevotellaceae"
prevo = data.frame( model = model, prevo )
prevo = as_tibble( prevo )

# Rikenellaceae
conf.int = data.frame( confint( glmm.int.Rikenellaceae ))
colnames( conf.int )[ 1 ] = "conf.low"
colnames( conf.int )[ 2 ] = "conf.high"
conf.int = conf.int[ -c( 11:16 ), ]

rikenell = broom.mixed::tidy( glmm.int.Rikenellaceae )[ , -c( 1 )]
rikenell = rikenell[ -c( 11:18 ), ]
rikenell = rikenell[ , -c( 1 )]
effect = "fixed"

rikenell = data.frame( model = "Rikenellaceae", effect = effect, rikenell, conf.low = conf.int$conf.low, conf.high = conf.int$conf.high )
rikenell = rikenell[ , -c( 8 )]
rikenell = as_tibble( rikenell )

# Ruminococcaceae
rumino = broom.mixed::tidy( lme.int.Ruminococcaceae, conf.int = TRUE )
rumino = rumino[ -c( 11:12 ), -c( 6,8 )]
model = "Ruminococcaceae"
rumino = data.frame( model = model, rumino )
rumino = as_tibble( rumino )

# Sutterellaceae

# Veillonellaceae
conf.int = data.frame( confint( glmm.slope.Veillonellaceae ))
colnames( conf.int )[ 1 ] = "conf.low"
colnames( conf.int )[ 2 ] = "conf.high"
conf.int = conf.int[ -c( 11:17 ), ]

veillon = broom.mixed::tidy( glmm.slope.Veillonellaceae )[ , -c( 1 )]
veillon = veillon[ -c( 11:18 ), ]
veillon = veillon[ , -c( 1 )]
effect = "fixed"

veillon = data.frame( model = "Veillonellaceae", effect = effect, veillon, conf.low = conf.int$conf.low, conf.high = conf.int$conf.high )
veillon = veillon[ , -c( 8 )]
veillon = as_tibble( veillon )
```


```{r Coefficient Plot, echo=TRUE, fig.width = 9, fig.height = 9}
# combine
common <- intersect( colnames( prevo ), colnames( rumino ))
all_bacteria = rbind( bacteroid[ common ], clostrid[ common ], clostridfam[ common ], prevo[ common ], lachno[ common ], rumino[ common ], 
                      rikenell[ common ], oscillo[ common ], eggerth[ common ], veillon[ common ] )

# pdf( "coefficientsplot.pdf", width = 15, height = 11 )
# dwplot( all_bacteria, dodge_size = 0.8, dot_args = list( size = 3 ),
#   whisker_args = list( size = 2 )) + xlab( "Coefficient Estimate" ) + geom_vline( xintercept = 0, colour = "grey60", linetype = 2 ) + theme_bw( base_size = 4 ) + theme( legend.title = element_blank(), text = element_text( size = 20 ))

dwplot( all_bacteria ) + xlab( "Coefficient Estimate" ) + geom_vline( xintercept = 0, colour = "grey60", linetype = 2 ) + theme_bw() + theme( legend.title = element_blank())
# dev.off()
```
